---
title: "Small granivores originals"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rats <- read.csv(here::here("lore", "1994_longterm", "1994_data_complete.csv"), stringsAsFactors = F)

rat_plot_totals <- read.csv(here::here("lore", "1994_longterm", "1994_data_plot_totals.csv"), stringsAsFactors = F)
library(dplyr)
library(ggplot2)

```

```{r}

rat_plot_totals <- rat_plot_totals %>%
 filter(grepl("orig", plot_type)) %>%
  mutate(oplot_type = ordered(plot_type),
         oplot = ordered(plot))
```

## 1. Did the small granivore response on the **original** exclosure plots hold up over time?

Comparing original exclosures to original controls.

This is the same analysis as for the 1981 paper, but with additional data.

I have split the timeseries for all plots into 2 sections. The first is 1977-85. The second is 85-91. This is chosen so the second chunk consists of 3 years pre treatment and 3 years post. It might make sense to have a shorter pre-treatment section in the second chunk.

```{r}
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))

sg_orig <- filter(rat_plot_totals, type == "small_granivore", period < 35)

ggplot(sg_orig, aes(period, nind, color = oplot_type, group = plot)) +
  geom_line(size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) + 
  ggtitle("Section 1, 1977-1985") +
  theme(legend.position = "top")

library(mgcv)

sg.gam <- gam(nind ~ oplot_type + s(period) + s(period, by = oplot_type) + oplot + s(period, by = oplot), data = sg_orig, family = poisson, method = "REML")

summary(sg.gam)

gam.check(sg.gam)

sg.pdat <- make_pdat(sg_orig, comparison_variable = "oplot_type", include_plot = T)
sg.pred <- get_predicted_vals(sg.gam, sg.pdat)

sg.diff <- get_exclosure_diff(sg.gam, sg.pdat,comparison_variable = "oplot_type")

print(plot_exclosure_diff(sg.diff) +
        geom_vline(xintercept = 35))

sg.pred <- add_exclosure_diff(sg.pred, sg.diff)

print(plot_link_pred(sg.pred, comparison_variable = "oplot_type") +
        geom_vline(xintercept = 35))

print(plot_fitted_pred(sg.pred, comparison_variable = "oplot_type") +
        geom_vline(xintercept = 35))

min(sg.diff$period[ which(!sg.diff$diff_overlaps_zero)])
```

(vertical line is where the 1981 data stops.)

Yep.

Some notes.

- Note `k=25` for the GAM; this is to get the k-index over 1. 
- The *timing* of the response is remarkably consistent (still being placed at 12-13).
- The *magnitude* of the response fluctuates over time but often far exceeds what was seen in the first 35 periods. 
- Still not sure about including plot as an effect.


