---
title: "to 2015"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)

```

```{r}
rats <- read.csv(here::here("lore", "2020_redux", "2020_data_complete.csv"))

```

```{r}

rats_types_totals <- rats %>%
  mutate(plot = factor(plot),
         treatment = factor(treatment)) %>%
  #mutate(plot = as.ordered(plot),
 #        treatment = as.ordered(treatment)) %>%
  select(period, treatment, type, plot, energy) %>%
  group_by(period, treatment, type, plot) %>%
  summarize(nind = dplyr::n(),
            totale = sum(energy))  %>%
  ungroup() 
```

```{r}
ggplot(filter(rats_types_totals, type == "dipo", type != "other", treatment != "orig_exclosure"), aes(period, nind, color = treatment, group = plot)) +
 geom_smooth(method = "gam", se = F, size = 2) +  theme_bw() +
  scale_color_viridis_d(end = .8) 



ggplot(filter(rats_types_totals, type == "small_granivore", type != "other"), aes(period, nind, group = plot, color = treatment)) +
  geom_smooth(method = "gam", se = F, size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) 


```

```{r}

sg <- filter(rats_types_totals, type == "small_granivore") %>%
  mutate(oplot = ordered(plot, levels = c(4,2,3,8,11,14,15,17,18,21,22)),
         treatment = as.ordered(treatment)) %>%
  as.data.frame()


library(mgcv)
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))

#sg.mod <- gam(nind ~  treatment + s(period, k = 70) + s(period, by = treatment, k = 70), family = "poisson", data  = sg)

sg.mod <- gam(nind ~  treatment + s(period, k = 10) + s(period, by = treatment, k = 10), family = "poisson", data  = sg)

summary(sg.mod)

gam.check(sg.mod)

sg.pdat <- make_pdat(sg, include_plot = F)

sg.pred <- get_predicted_vals(sg.mod, sg.pdat)

```

Here we have a GAM - fit without plot, because when I included plot I got super-wrong estimates (like infinity rats for orig_exclosure????)

This GAM is trash according to gam.check.

```{r}

plot_fitted_pred(sg.pred)# + facet_wrap(vars(compare_var), scales = "free_y")

```

At this stage, the new_exclosure plots are still controls. Interestingly, they diverge from the always-control plots.

Also note the overall increase in the number of small granivores across the site, on exclosure and control plots. 

And that the differences (in the number of individuals of smgran species) actually seem less pronounced now than at the beginning of the study. 


```{r}

ctrl_v_orig <- get_exclosure_diff(sg.mod, sg.pdat, comparison_level = 3)

plot_exclosure_diff(ctrl_v_orig)

sg.origdiff <- add_exclosure_diff(sg.pred, ctrl_v_orig)

plot_fitted_pred(filter(sg.origdiff, treatment != "new_exclosure"))

```


So I'd caution against over-interpreting this, because there are issues with the GAM (gam.check suggests k is too low, but fitting higher k takes a long time and sometimes gets into weird spaces where the fitted values are not remotely realistic). However, it's suggestive that the difference in the **number of small granivores** on control v. exclosure plots has declined over time. 


Assuming this result held - would there be uestions then, whether the strength of the competitive effect of dipo on small granivores has weakened? This is one way of measuring that effect, but I'm not sure. There's the finding (from Erica's paper) that the reintroduction of krats made the small granivores "flee" - although on digging, I don't (right away) find evidence of that? E's paper is about how the treatment plots compare to controls once they are switched back to controls. Ellen's paper is actually that PP increases its preference for CONTROLS over EXCLOSURES after PB shows up.


```{r}

ctrl_v_new <- get_exclosure_diff(sg.mod, sg.pdat)

plot_exclosure_diff(ctrl_v_new)

sg.newdiff <- add_exclosure_diff(sg.pred, ctrl_v_new)

plot_fitted_pred(filter(sg.newdiff, treatment != "orig_exclosure"))

```


```{r}


plot_fitted_pred(filter(sg.origdiff, treatment != "new_exclosure", period > 400))


```