---
title: "New plots with 81 data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rats <- read.csv(here::here("lore", "1981_competition", "1981_data_complete.csv"), stringsAsFactors = F)
rat_totals <- read.csv(here::here("lore", "1981_competition", "1981_data_statevars.csv"), stringsAsFactors = F)
library(dplyr)
library(ggplot2)
library(gratia)
```

```{r, echo = F}

ggplot(filter(rat_totals, type != "other"), aes(period, nind, color = brown_trtmnt)) +
  geom_line() +
  geom_point()+
  theme_bw() +
  facet_wrap(vars(type), ncol = 1, scales = "free_y") +
  geom_vline(xintercept = 3.5) +
  geom_hline(yintercept = 0) +
  scale_colour_viridis_d(end = .8)

```
```{r}

rat_totals <- rat_totals 

load_mgcv()

rat_totals <- rat_totals %>%
  mutate(brown_trtmnt = as.factor(brown_trtmnt))

sg <- filter(rat_totals, type == "small_granivore")

n_gam <- gam(nind ~  brown_trtmnt + s(period, by = brown_trtmnt), data = sg, method = "REML", family = "poisson")

draw(n_gam)

n_gam_fitted <- add_fitted(sg, n_gam, value = "fitted")

ggplot(n_gam_fitted, aes(period, nind, color = brown_trtmnt)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  geom_line(aes(period, fitted, color = brown_trtmnt), size = 2) +
  scale_color_viridis_d(end = .8)

fitted_ci <- function(gam_obj, ndraws = 500, df, seed = 1977) {
  
  sampled_vals <- fitted_samples(gam_obj, n = ndraws, newdata = df, seed = seed)
  
  sampled_vals <- sampled_vals %>%
    group_by(row) %>%
    summarize(
      meanfit = mean(fitted),
      lowerfit = quantile(fitted, probs = .025),
      upperfit= quantile(fitted, probs = .975)
    ) %>%
    ungroup()
  
  df <- df %>%
    mutate(row = dplyr::row_number()) %>%
    left_join(sampled_vals)
  
  df  
}

n_gam_ci_manual <- fitted_ci(n_gam, df= sg)


ggplot(n_gam_ci_manual, aes(period, nind, color = brown_trtmnt)) +
  geom_line() +
  geom_line(aes(period, meanfit)) +
  geom_ribbon(aes(period, ymin = lowerfit, ymax = upperfit, fill = brown_trtmnt), alpha = .5) +
  theme_bw() +
  theme(legend.position = "top")


n_gam_diff <- difference_smooths(n_gam, smooth = "s(period)")

ggplot(n_gam_diff, aes(period, diff)) +
  geom_line() +
  geom_ribbon(aes(period, ymin = lower,ymax= upper), alpha = .5) +
  geom_hline(yintercept = 0)

```


This is fitting a single GAM to the small granivores by treatment. There is no plot effect included - I haven't checked to see if there are differences between plots. So this is rough.

The difference_smooths plot kind of matches but doesn't completely track with the estimates/fitted CI I generated via `gratia::fitted_samples`. The samples appear to overlap at the beginning, when the difference plot would have that the dipo_present treatment is higher than the dipo_absent treatment (although note that those first 3-5 periods are right as the treatments were being implemented). Then, around period 30, the samples appear divergent while the difference plot has the difference as much closer to 0 than it appears. 


```{r}
load_mgcv()

rat_totals <- rat_totals %>%
  mutate(brown_trtmnt = as.factor(brown_trtmnt))

so <- filter(rat_totals, type == "small_omnivore")

n_gam <- gam(nind ~  brown_trtmnt + s(period, by = brown_trtmnt), data = so, method = "REML", family = "poisson")

draw(n_gam)

n_gam_fitted <- add_fitted(so, n_gam, value = "fitted")

ggplot(n_gam_fitted, aes(period, nind, color = brown_trtmnt)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  geom_line(aes(period, fitted, color = brown_trtmnt), size = 2) +
  scale_color_viridis_d(end = .8)

fitted_ci <- function(gam_obj, ndraws = 500, df, seed = 1977) {
  
  sampled_vals <- fitted_samples(gam_obj, n = ndraws, newdata = df, seed = seed)
  
  sampled_vals <- sampled_vals %>%
    group_by(row) %>%
    summarize(
      meanfit = mean(fitted),
      lowerfit = quantile(fitted, probs = .025),
      upperfit= quantile(fitted, probs = .975)
    ) %>%
    ungroup()
  
  df <- df %>%
    mutate(row = dplyr::row_number()) %>%
    left_join(sampled_vals)
  
  df  
}

n_gam_ci_manual <- fitted_ci(n_gam, df= so)


ggplot(n_gam_ci_manual, aes(period, nind, color = brown_trtmnt)) +
  geom_line() +
  geom_line(aes(period, meanfit)) +
  geom_ribbon(aes(period, ymin = lowerfit, ymax = upperfit, fill = brown_trtmnt), alpha = .5) +
  theme_bw() +
  theme(legend.position = "top")


n_gam_diff <- difference_smooths(n_gam, smooth = "s(period)")

ggplot(n_gam_diff, aes(period, diff)) +
  geom_line() +
  geom_ribbon(aes(period, ymin = lower,ymax= upper), alpha = .5) +
  geom_hline(yintercept = 0)

```


```{r}
load_mgcv()

rat_totals <- rat_totals %>%
  mutate(brown_trtmnt = as.factor(brown_trtmnt))

dipo <- filter(rat_totals, type == "dipo")

n_gam <- gam(nind ~  brown_trtmnt + s(period, by = brown_trtmnt), data = dipo, method = "REML", family = "poisson")

draw(n_gam)

n_gam_fitted <- add_fitted(dipo, n_gam, value = "fitted")

ggplot(n_gam_fitted, aes(period, nind, color = brown_trtmnt)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  geom_line(aes(period, fitted, color = brown_trtmnt), size = 2) +
  scale_color_viridis_d(end = .8)

fitted_ci <- function(gam_obj, ndraws = 500, df, seed = 1977) {
  
  sampled_vals <- fitted_samples(gam_obj, n = ndraws, newdata = df, seed = seed)
  
  sampled_vals <- sampled_vals %>%
    group_by(row) %>%
    summarize(
      meanfit = mean(fitted),
      lowerfit = quantile(fitted, probs = .025),
      upperfit= quantile(fitted, probs = .975)
    ) %>%
    ungroup()
  
  df <- df %>%
    mutate(row = dplyr::row_number()) %>%
    left_join(sampled_vals)
  
  df  
}

n_gam_ci_manual <- fitted_ci(n_gam, df= dipo)


ggplot(n_gam_ci_manual, aes(period, nind, color = brown_trtmnt)) +
  geom_line() +
  geom_line(aes(period, meanfit)) +
  geom_ribbon(aes(period, ymin = lowerfit, ymax = upperfit, fill = brown_trtmnt), alpha = .5) +
  theme_bw() +
  theme(legend.position = "top")


n_gam_diff <- difference_smooths(n_gam, smooth = "s(period)")

ggplot(n_gam_diff, aes(period, diff)) +
  geom_line() +
  geom_ribbon(aes(period, ymin = lower,ymax= upper), alpha = .5) +
  geom_hline(yintercept = 0)

```



I feel like there might be something I'm not grasping here around the parametric term? Shouldn't the dipo_absent **always** be lower than dipo_present (for the krats on the krat removals)? Unless the difference smooths is showing the difference **without taking into account the _partial_ effect of treatment**