---
title: "New plots with 81 data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rats <- read.csv(here::here("lore", "1981_competition", "1981_data_complete.csv"), stringsAsFactors = F)
rat_totals <- read.csv(here::here("lore", "1981_competition", "1981_data_statevars.csv"), stringsAsFactors = F)
library(dplyr)
library(mgcv)
library(ggplot2)
```

```{r, echo = F}

ggplot(filter(rat_totals, type != "other"), aes(period, nind, color = brown_trtmnt)) +
  geom_line() +
  geom_point()+
  theme_bw() +
  facet_wrap(vars(type), ncol = 1, scales = "free_y") +
  geom_vline(xintercept = 3.5) +
  geom_hline(yintercept = 0) +
  scale_colour_viridis_d(end = .8)

```

## Analyzed using GAMs

Following Christensen et al (2019), we can use the *difference in smooths* fitted using a GAM as a way to tell if and when two timeseries are diverging/converging. 

### Difference smooths walkthrough

Here is a walkthrough (fairly detailed) of how we obtain the difference in smooths. This approach is modeled after the analysis in Christensen et al 2019 (code in Erica's github and also copied to this repo in `lore/2019_switch`) and blog posts by Gavin Simpson. I have re-written the code mostly for my own understanding; bits of it probably echo Erica and Gavin closely.

#### Fitting the GAM

- (For now at least) we're working with count data, so the link is Poisson. This means we get the differences in the difference smooth on the log scale.
- I have fit using **treatment** but not **plot**. It may be best practice to include terms for plot (see Christensen et al 2019) but that would complicate the machinery for now. 
- I am using the default `k = 10` for the number of basis functions.
- I am fitting using the formula `count ~ okrat_treatment + s(period) + s(period, by  okrat_treatment)`, where `okrat_treatment` is an ordered factor with levels for control and krat exclosure. This means:
    - `okrat_treatment`: There is an intercept like term for the *difference in means* between treatments. Because `okrat_treatment` is an ordered factor, they aren't really intercepts, they're weird exponential things.
    - `s(period)`: This fits the smooth for the *reference level* of `okrat_treatment`, in this case `control`.
    - `s(period, by = okrat_treatment)`: This fits the smooth for the *difference* between the remaining levels of `okrat_treatment` and the reference level. In this case there is only one other level.
    
```{r}

rat_totals <- rat_totals %>%
  mutate(krat_treatment = ifelse(brown_trtmnt == "dipo_present", "control", "exclosure")) %>%
  mutate(okrat_treatment = ordered(krat_treatment))

sg <- filter(rat_totals, type == "small_granivore")

ggplot(sg, aes(period, nind, color= okrat_treatment)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  scale_color_viridis_d(end = .8)
```
```{r}

nind.gam <- gam(nind ~ okrat_treatment + s(period) + s(period, by = okrat_treatment), data = sg, family = poisson, method = "REML")

plot(nind.gam, pages= 1, scale = 0)


```

These are the default plots from `mgcv` for a GAM fit with an ordered factor. The plot on the left is the fit for the reference level of the factor, so control. The plot on the right is the adjustment between the reference level and the second level, exclosure.


We can extract the predicted values and a SE envelope for these smooths. First on the link (log) scale, and then invlink-ed back to count values.

The control smooth on the link scale matches the left-side plot from `mgcv`. The exclosure smooths should be (about) the control smooth plus the difference.

```{r}

pdat <- as.data.frame(expand.grid(period = seq(min(sg$period), max(sg$period), length.out= 500), okrat_treatment = levels(sg$okrat_treatment)))

nind.predicted <- predict(nind.gam, newdata = pdat, type = "lpmatrix")
nind.link <- predict(nind.gam, newdata = pdat, type = "link", se.fit = T)

nind.predicted.vals <- nind.gam$family$linkinv(nind.predicted %*% coefficients(nind.gam))

pdat.pred <- pdat %>%
  mutate(predicted = nind.predicted.vals,
         link = nind.link$fit,
         se_link = nind.link$se.fit) %>%
  mutate(invlink_fit = nind.gam$family$linkinv(link),
         invlink_upper = nind.gam$family$linkinv(link + (2 * se_link)),
         invlink_lower = nind.gam$family$linkinv(link - (2 * se_link)),
         link_upper = link + (2 * se_link),
         link_lower = link - (2 * se_link))


ggplot(pdat.pred, aes(period, link, color= okrat_treatment)) +
  geom_line() +
  geom_ribbon(aes(period, ymin = link_lower, ymax = link_upper, fill = okrat_treatment), alpha = .5) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  scale_fill_viridis_d(end = .8)

ggplot(pdat.pred, aes(period, invlink_fit, color= okrat_treatment)) +
  geom_line() +
  geom_ribbon(aes(period, ymin = invlink_lower, ymax = invlink_upper, fill = okrat_treatment), alpha = .5) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  scale_fill_viridis_d(end = .8)

```

We can then calculate the *difference* between these smooths, plus a credible interval. This then allows us to find when the difference is overlapping 0 (no treatment effect) and when it is not. 

We do this keeping **all** the terms. That is, including the quasi-intercept to account for the difference in means between treatments. Erica's models included terms for plot, which she then excluded from calculating the differences. That's not in play here (yet?).

```{r}


nind.diff.keeppar <- nind.predicted[1:500, ] - nind.predicted[501:1000, ]

nind.diff.vals <- nind.diff.keeppar %*% coef(nind.gam)


nind.diff.se<- sqrt(rowSums((nind.diff.keeppar %*% vcov(nind.gam, unconditional = FALSE)) * nind.diff.keeppar))

crit <- qnorm(.05/2, lower.tail = FALSE)
upr <- nind.diff.vals + (crit * nind.diff.se)
lwr <- nind.diff.vals - (crit * nind.diff.se)


pdat.diff <- pdat %>%
  select(period) %>%
  distinct()%>%
  mutate(fitted_dif = nind.diff.vals,
         upper= upr,
         lower = lwr) %>%
  mutate(diff_overlaps_zero = (upper * lower) < 0
  )

ggplot(pdat.diff, aes(period, fitted_dif)) +
  geom_line() +
  geom_ribbon(aes(period, ymin = lower, ymax = upper), alpha  = .5) +
  geom_hline(yintercept = 0) +
  geom_point(data = filter(pdat.diff, diff_overlaps_zero), aes(period, 1), color  = "red", size = 2) +
  theme_bw()

```

The red line shows where the difference between the two smooths' CI is overlapping 0; i.e., we aren't really confident that the two smooths are different.

```{r}

pdat.pred <- left_join(pdat.pred, select(pdat.diff, period, diff_overlaps_zero))


ggplot(pdat.pred, aes(period, link, color= okrat_treatment)) +
  geom_line() +
  geom_ribbon(aes(period, ymin = link_lower, ymax = link_upper, fill = okrat_treatment), alpha = .5) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  scale_fill_viridis_d(end = .8) +
  geom_point(data = filter(pdat.pred, diff_overlaps_zero), aes(period, 0), color = "red")


ggplot(pdat.pred, aes(period, invlink_fit, color= okrat_treatment)) +
  geom_line() +
  geom_ribbon(aes(period, ymin = invlink_lower, ymax = invlink_upper, fill = okrat_treatment), alpha = .5) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  scale_fill_viridis_d(end = .8) +
  geom_point(data = filter(pdat.pred, diff_overlaps_zero), aes(period, 0), color = "red")
```


This doesn't quite match what we would think from the `mgcv` plot, I think because the `mgcv` plot is showing the difference smooth without the intercept adjustment.

I re-did this but swapped in Erica's data and you get the results that match the 2019 paper. So I think it is basically sound? 

What gives me the creeps is that the overlapping-zero interval 1) differs from the mgcv plot and 2) doesn't match what you'd think if you just looked at the plots of the smooths themselves.

However, 1) may be because the mgcv plot doesn't include the offset for different treatment means. You see this especially in Erica's data, where the mgcv plot for exclosure only briefly overlaps 0 - because of the huge offset, maybe? Anyway, this arithmetic for calculating the difference smooths, including the offset term, qualitatively matches the paper (even though there might be minor differences because the paper included a plot term and had different k's). It also matches in 2), in that the *smooths* appear to overlap at times when the *difference smooth* does not overlap 0 (e.g. the very end of the TS for the total exclosure v. control plots, the very last plot in gams_switchdata.Rmda).


### Interpreting

```{r}

min(filter(pdat.pred, period > 4, !diff_overlaps_zero)$period)

```

So this analysis actually picks up a difference in small granivore abundance on krat exclosures vs. control plots as early as period 13-14, or about 10 months after the treatments were implemented. 


The paper identifies the change as starting in July 1978, or "eight months".

```{r}

filter(rat_totals, period == 13) %>%
  select(period, year, month)%>%
  distinct()

```

OMG. O M G.

<!-- ## removing intercept -->


<!-- ```{r} -->

<!-- nind.predicted[,2] <- 0 -->

<!-- nind.diff.keeppar <- nind.predicted[1:500, ] - nind.predicted[501:1000, ] -->

<!-- nind.diff.vals <- nind.diff.keeppar %*% coef(nind.gam) -->


<!-- nind.diff.se<- sqrt(rowSums((nind.diff.keeppar %*% vcov(nind.gam, unconditional = FALSE)) * nind.diff.keeppar)) -->

<!-- crit <- qnorm(.05/2, lower.tail = FALSE) -->
<!-- upr <- nind.diff.vals + (crit * nind.diff.se) -->
<!-- lwr <- nind.diff.vals - (crit * nind.diff.se) -->


<!-- pdat.diff <- pdat %>% -->
<!--   select(period) %>% -->
<!--   distinct()%>% -->
<!--   mutate(fitted_dif = nind.diff.vals, -->
<!--          upper= upr, -->
<!--          lower = lwr) %>% -->
<!--   mutate(diff_overlaps_zero = (upper * lower) < 0 -->
<!--   ) -->

<!-- ggplot(pdat.diff, aes(period, fitted_dif)) + -->
<!--   geom_line() + -->
<!--   geom_ribbon(aes(period, ymin = lower, ymax = upper), alpha  = .5) + -->
<!--   geom_hline(yintercept = 0) + -->
<!--   geom_point(data = filter(pdat.diff, diff_overlaps_zero), aes(period, 1), color  = "red", size = 2) + -->
<!--   theme_bw() -->

<!-- ``` -->

<!-- The red line shows where the difference between the two smooths' CI is overlapping 0; i.e., we aren't really confident that the two smooths are different. -->

<!-- ```{r} -->

<!-- pdat.pred <- left_join(select(pdat.pred, -diff_overlaps_zero), select(pdat.diff, period, diff_overlaps_zero)) -->



<!-- ggplot(pdat.pred, aes(period, invlink_fit, color= okrat_treatment)) + -->
<!--   geom_line() + -->
<!--   geom_ribbon(aes(period, ymin = invlink_lower, ymax = invlink_upper, fill = okrat_treatment), alpha = .5) + -->
<!--   theme_bw() + -->
<!--   scale_color_viridis_d(end = .8) + -->
<!--   scale_fill_viridis_d(end = .8) + -->
<!--   geom_point(data = filter(pdat.pred, diff_overlaps_zero), aes(period, 0), color = "red") -->
<!-- ``` -->


