---
title: "to 2015"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)

```

```{r}
rats <- read.csv(here::here("lore", "2020_redux", "2020_data_complete.csv"))

```

```{r}

smgran <- filter(rats, type == "small_granivore")

smgran_comp <- smgran %>%
  select(censusdate, treatment, plot, species) %>%
  mutate(censusyear = as.integer(substr(censusdate, 0, 4))) %>%
  group_by( treatment, censusyear, plot, species) %>%
  summarize(nind = dplyr::n()) %>%
  ungroup %>%
  group_by(censusyear, treatment, plot) %>%
  mutate(totalind = sum(nind)) %>%
  ungroup() %>%
  mutate(nind = nind / totalind) %>% tidyr::pivot_wider(names_from = species, values_from = nind, values_fill= 0)



smgran_comp_periods <- smgran %>%
  select(censusdate, period, treatment, plot, species) %>%
  group_by( censusdate, period, treatment, plot, species) %>%
  summarize(nind = dplyr::n()) %>%
  ungroup %>%
  group_by(censusdate, period, treatment, plot) %>%
  mutate(totalind = sum(nind)) %>%
  ungroup() %>%
  mutate(nind = nind / totalind) %>% tidyr::pivot_wider(names_from = species, values_from = nind, values_fill= 0) %>%
  mutate(censusdate = as.Date(censusdate))

```

```{r}

ggplot(smgran_comp, aes(censusyear, PP, group = plot)) +
  geom_line(size = 2, color = "red", alpha= .5) +
  geom_line(aes(censusyear, PB, group = plot), size = 2, color = "blue", alpha = .5) +
  theme_bw() +
  facet_wrap(vars(plot))



ggplot(smgran_comp, aes(censusyear, PP + PB, group = plot)) +
  geom_line(size = 2, color = "red", alpha= .5)  +
  theme_bw() +
  facet_wrap(vars(treatment))
```


```{r}


ggplot(smgran_comp_periods, aes(censusdate, PP + PB)) +
  geom_smooth(method = "gam", se = T, size = 2, color = "red", alpha= .5)  +
   geom_smooth(aes(censusdate, PP), method = "gam", se = T, size = 2, color = "blue", alpha= .5)  +
  theme_bw() +
  facet_wrap(vars(treatment), ncol = 1) +
  scale_x_date(date_breaks = "9 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))

```

This is just to say...

* While this wasn't the case early in the study, after the late 1990s, PP + PB account for most of the "small granivore" community. 
* After around 2010, most (but not all) of it is PP. 
* PPs were a relatively small proportion even of **small granivores** in the 80s and early 90s. 

Generally - possibly - pointing to the argument that, for the new manipulations, we are looking a story that is perhaps mostly a story about how PP, PB, and DX play together...

Bailey's might be fading out for reasons unknown to us.

Where Bailey's is present, especially since the 2010s, it's not very abundant. 

See Ellen's paper for some on PB-PP interactions, also Thibault et al Redundant. 

If we say PPs would rather hang with krats than with PB,

and krats are having any kind of suppressive effects on either or both of them,

removing krats might create a kind of vacuum into which either PP or PB could expand.

we know in the past, removing krats led to SOME expansion of pp's/smgran (70s and 80s) and that, once they found the site, bailey's preferentially colonized exclosures over controls.

however, since then we have seen the shift to shrubland.

krats may no longer be being very suppressive?