---
title: "Energy plots"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.dim = c(7, 3))
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
source(here::here("scaffold", "R", "data.R"))
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))
library(mgcv)

```

variables...

- Dipo energy on a control plot (amount of energy left available given Dipo removal)
- tinygran energy on a control plot (baseline tinygran energy if no compensation happens)
- Additional tinygran energy on a treatment plot (amount of energy attributable to compensation)
- tinygran_gain / dipo_control (approximates amount of POTENTIAL compensation ACHIEVED, but can be negative or > 1) ---> this is what we want to estimate.

```{r}


plot_totals <- get_rodent_data(use_christensen_plots = F, return_plot = T) %>%
   mutate(tinygran_e = smgran_e - pb_e) %>%
  mutate(oplottype = ordered(plot_type))

controls <- filter(plot_totals, oplottype == "CC")
# 
# control_gam <- gam(total_e ~ s(period), family = "tw", data = controls)
# 
# control_samples <- gratia::fitted_samples(control_gam, n = 100, seed = 1977, scale = "response", newdata = data.frame(period = unique(controls$period)))
# 
# control_samples <- control_samples %>%
#   left_join(data.frame(period = unique(controls$period), row = 1:length(unique(controls$period)))) %>%
#   mutate(draw = as.factor(draw))
# 
# ggplot(control_samples, aes(period, fitted, group = draw)) +
#   geom_line(alpha = .1)



dipo_control_gam <- gam(dipo_e ~ s(period), family = "tw", data = controls)

dipo_control_samples <- gratia::fitted_samples(dipo_control_gam, n = 100, seed = 1977, scale = "response", newdata = data.frame(period = unique(controls$period)))

dipo_control_samples <- dipo_control_samples %>%
  left_join(data.frame(period = unique(controls$period), row = 1:length(unique(controls$period)))) %>%
  mutate(draw = as.factor(draw))

ggplot(dipo_control_samples, aes(period, fitted, group = draw)) +
  geom_line(alpha = .1)


tinygran_control_gam <- gam(tinygran_e ~ s(period), family = "tw", data = controls)

tinygran_control_samples <- gratia::fitted_samples(tinygran_control_gam, n = 100, seed = 1977, scale = "response", newdata = data.frame(period = unique(controls$period)))

tinygran_control_samples <- tinygran_control_samples %>%
  left_join(data.frame(period = unique(controls$period), row = 1:length(unique(controls$period)))) %>%
  mutate(draw = as.factor(draw))

ggplot(tinygran_control_samples, aes(period, fitted, group = draw)) +
  geom_line(alpha = .1)


exclosures <- filter(plot_totals, oplottype == "CE")


tinygran_exclosure_gam <- gam(tinygran_e ~ s(period), family = "tw", data = exclosures)

tinygran_exclosure_samples <- gratia::fitted_samples(tinygran_exclosure_gam, n = 100, seed = 1977, scale = "response", newdata = data.frame(period = unique(exclosures$period)))

tinygran_exclosure_samples <- tinygran_exclosure_samples %>%
  left_join(data.frame(period = unique(exclosures$period), row = 1:length(unique(exclosures$period)))) %>%
  mutate(draw = as.factor(draw))

ggplot(tinygran_exclosure_samples, aes(period, fitted, group = draw)) +
  geom_line(alpha = .1) +
  geom_line(data = tinygran_control_samples, color = "blue", alpha = .1)


tinygran_sampled_diff <- tinygran_exclosure_samples %>%
  mutate(diff = fitted - tinygran_control_samples$fitted)

ggplot(tinygran_sampled_diff, aes(period, diff, group = draw)) +
  geom_line(alpha = .1) 

tinygran_sampled_prop <- tinygran_sampled_diff %>%
  mutate(prop = diff / dipo_control_samples$fitted)

ggplot(tinygran_sampled_prop, aes(period, prop, group = draw)) +
  geom_line(alpha = .1) 

tinygran_sampled_prop_era <- left_join(tinygran_sampled_prop, distinct(select(plot_totals, period, era, censusdate)))

ggplot(tinygran_sampled_prop_era, aes(censusdate, prop, group = draw)) +
  geom_line(alpha = .1) +
  facet_grid(cols = vars(era), scales = "free_x", space = "free")


```

```{r}

all_together <- controls %>%
  select(period, censusdate, era, dipo_e, tinygran_e, plot) %>%
  tidyr::pivot_longer(c(dipo_e, tinygran_e), names_to = "rods", values_to = "e") %>%
  mutate(rods = paste0(rods, "_c")) %>%
  bind_rows(mutate(rename(select(exclosures, period, censusdate, era, plot, tinygran_e), e = tinygran_e), rods = "tinygran_e_e")) %>%
  mutate(frods = as.factor(rods),
         orods = ordered(frods)) %>%
  mutate(rod_group = ifelse(grepl("tinygran", rods), "tinygran", "dipo")) %>%
  mutate(rod_group = as.factor(rod_group))

all_together_gam <- gam(e ~ s(period, by = orods, k = 100) + orods + s(period, k = 100) + s(plot, bs = "re"), family = "tw", data = all_together)


all_together_pdat <- select(all_together, period, orods, rod_group) %>%
  distinct() %>%
  mutate(row = dplyr::row_number())  %>%
  mutate(plot = 4)

all_together_samples <- gratia::fitted_samples(all_together_gam, n = 100, seed = 1977, scale = "response", newdata = all_together_pdat, exclude = "s(plot)")

all_together_samples <- all_together_samples %>%
  left_join(all_together_pdat) %>%
  mutate(draw = as.factor(draw))

ggplot(all_together_samples, aes(period, fitted, group = draw, color = orods)) +
  geom_line(alpha = .1) +
  geom_point(data = filter(all_together_samples, period > 1000)) +
  facet_wrap(vars(orods))

all_together_xp <-   predict(all_together_gam, newdata = all_together_pdat, type = 'lpmatrix', exclude = "s(plot)")

dipo_xp <- all_together_xp[ filter(all_together_pdat, orods == "dipo_e_c")$row, ]
sg_c_xp <- all_together_xp[ filter(all_together_pdat, orods == "tinygran_e_c")$row, ]
sg_e_xp <- all_together_xp[ filter(all_together_pdat, orods == "tinygran_e_e")$row, ]

dipo_cols <- colnames(all_together_xp)[ which(grepl('dipo', colnames(all_together_xp)))]
sg_c_cols <- colnames(all_together_xp)[ which(grepl('tinygran_e_c', colnames(all_together_xp)))]
sg_e_cols <- colnames(all_together_xp)[ which(grepl('tinygran_e_e', colnames(all_together_xp)))]
alpha = .05
sg_e_minus_c_xp <- sg_e_xp - sg_c_xp
sg_e_minus_c_diff <- sg_e_minus_c_xp %*% coef(all_together_gam)
sg_e_minus_c_se <- sqrt(rowSums((sg_e_minus_c_xp %*% vcov(all_together_gam, unconditional = F)) * sg_e_minus_c_xp))
crit <- qnorm(alpha/2, lower.tail = FALSE)
sg_e_minus_c_upr <- sg_e_minus_c_diff + (crit * sg_e_minus_c_se)
sg_e_minus_c_lwr <- sg_e_minus_c_diff - (crit * sg_e_minus_c_se)
sg_e_minus_c_df <- data.frame(
             diff = sg_e_minus_c_diff,
             se = sg_e_minus_c_se,
             upper = sg_e_minus_c_upr,
             lower = sg_e_minus_c_lwr,
             period = unique(all_together_pdat$period),
             pair = "sg_e_minus_c") %>%
  mutate(diff_over_zero = upper * lower < 0)

ggplot(sg_e_minus_c_df, aes(period, diff)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5)

ilink <- all_together_gam$family$linkinv

# 
# sg_e_minus_c_minus_dipo_xp <- sg_e_minus_c_xp - dipo_xp
# sg_e_minus_c_minus_dipo_diff <- sg_e_minus_c_minus_dipo_xp %*% coef(all_together_gam)
# sg_e_minus_c_minus_dipo_se <- sqrt(rowSums((sg_e_minus_c_minus_dipo_xp %*% vcov(all_together_gam, unconditional = F)) * sg_e_minus_c_minus_dipo_xp))
# crit <- qnorm(alpha/2, lower.tail = FALSE)
# sg_e_minus_c_minus_dipo_upr <- sg_e_minus_c_minus_dipo_diff + (crit * sg_e_minus_c_minus_dipo_se)
# sg_e_minus_c_minus_dipo_lwr <- sg_e_minus_c_minus_dipo_diff - (crit * sg_e_minus_c_minus_dipo_se)
# sg_e_minus_c_minus_dipo_df <- data.frame(
#              diff = sg_e_minus_c_minus_dipo_diff,
#              se = sg_e_minus_c_minus_dipo_se,
#              upper = sg_e_minus_c_minus_dipo_upr,
#              lower = sg_e_minus_c_minus_dipo_lwr,
#              period = unique(all_together_pdat$period),
#              pair = "sg_minus_dipo")
# 
# ggplot(sg_e_minus_c_minus_dipo_df, aes(period, diff)) +
#   geom_line() +
#   geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5)

```


```{r}



ggplot(all_together_samples, aes(period, fitted, group = draw, color = orods)) +
  geom_line(alpha = .1) +
  geom_point(data = filter(all_together_samples, period > 1000)) +
  facet_wrap(vars(orods))

all_together_samples_wide <- all_together_samples %>%
  tidyr::pivot_wider(id_cols = c(draw, period), names_from = orods, values_from = fitted) %>%
  left_join(select(sg_e_minus_c_df, period, diff_over_zero))


ggplot(all_together_samples_wide, aes(period, tinygran_e_e, group = draw)) +
  geom_line(alpha = .1) +
  geom_line(aes(y = tinygran_e_c), color = "blue")
  geom_point(aes(period, 0, color = diff_over_zero))

ggplot(all_together_samples_wide, aes(period, tinygran_e_e - tinygran_e_c, group = draw)) +
  geom_line(alpha = .1) +
  geom_point(aes(period, 0, color = diff_over_zero))


ggplot(all_together_samples_wide, aes(period, (tinygran_e_e - tinygran_e_c)/dipo_e_c, group = draw)) +
  geom_line(alpha = .1) +
  geom_point(aes(period, 0, color = diff_over_zero))


all_together_samples_big <- gratia::fitted_samples(all_together_gam, n = 10000, seed = 1977, scale = "response", newdata = all_together_pdat, exclude = "s(plot)")%>%
  left_join(all_together_pdat) %>%
  mutate(draw = as.factor(draw))

all_together_samples_big_summary <- all_together_samples_big  %>%
  group_by(row, period, orods, rod_group) %>%
  summarize(meanfit = mean(fitted),
            lowerq = quantile(fitted, probs = .025),
            upperq = quantile(fitted, probs = .975)) %>%
  ungroup()

all_together_samples_big_summary_wide <- all_together_samples_big_summary %>%
  tidyr::pivot_wider(id_cols = c(period), names_from = orods, values_from = c(meanfit, lowerq, upperq)) %>%
  left_join(select(sg_e_minus_c_df, period, diff_over_zero))

all_together_samples_big_wide<- all_together_samples_big %>%
  tidyr::pivot_wider(id_cols = c(period, draw), names_from = orods, values_from = c(fitted)) %>%
  mutate(sg_e_minus_c = tinygran_e_e - tinygran_e_c,
         sg_e_minus_c_over_dipo = (tinygran_e_e - tinygran_e_c)/dipo_e_c)  %>%
  select(-dipo_e_c, -tinygran_e_c, -tinygran_e_e) %>%
  group_by(period) %>%
  summarize(mean_sg_minus_c = mean(sg_e_minus_c),
            lowerq_sg_minus_c = quantile(sg_e_minus_c, probs = .025),
            upperq_sg_minus_c = quantile(sg_e_minus_c, probs = .975),
            mean_sg_minus_c_over_dipo = mean(sg_e_minus_c_over_dipo),
            lowerq_sg_minus_c_over_dipo = quantile(sg_e_minus_c_over_dipo, probs = .025),
            upperq_sg_minus_c_over_dipo = quantile(sg_e_minus_c_over_dipo, probs = .975)) %>%
  ungroup()

all_together_samples_big_summary_wide <- all_together_samples_big_summary_wide %>%
  left_join(all_together_samples_big_wide)

ggplot(all_together_samples_big_summary_wide, aes(period, mean_sg_minus_c_over_dipo)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowerq_sg_minus_c_over_dipo, ymax = upperq_sg_minus_c_over_dipo), alpha = .1)+
  geom_point(aes(period, 0, color = diff_over_zero))


```