---
title: "GAM"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.dim = c(7, 3))
library(dplyr)
library(ggplot2)
library(mgcv)
theme_set(theme_bw())
source(here::here("scaffold", "R", "data.R"))
source(here::here("lore", "2019_switch", "FinalAnalysis", "analysis_functions.R"))

load(here::here("randomplots_gams.RData"))

get_smooth_diffs <- function(MODEL, dat) {
  
  cc_ce <- osmooth_diff(MODEL, dat, "numericdate", "CC", "CE", "oTreatment")
  cc_ee <- osmooth_diff(MODEL, dat, "numericdate", "CC", "EE", "oTreatment")
  
  diffs <- bind_rows(cc_ce, cc_ee) %>%
    mutate(diff_overlaps_zero = (upper * lower) < 0) %>%
    mutate(compare_trt = substr(pair, 4, 5)) 
  
  dat <- dat %>%
    mutate(compare_trt = as.character(oTreatment)) %>%
    left_join(diffs)
  
  return(dat)
}
```

### Small gran

```{r}

load(here::here("se_gams.RData"))

se_AICs <- lapply(list(se.gam, se.gam.noplot, se.gam.plotint, se.gam.randomplot, se.gam.randomplot.s), FUN = AIC)

exVars.d <- c('oPlot', paste0('s(numericdate):oPlot', unique(plot_totals$oPlot)))
preddat <- predict_treat_effect2(plot_totals, 500, se.gam, exVars.d) %>%
  select(censusdate, oPlot, oTreatment, numericdate)

se_diffs <- get_smooth_diffs(se.gam, preddat)
se_diffs_noplot <- get_smooth_diffs(se.gam.noplot, preddat)
se_diffs_random <- get_smooth_diffs(se.gam.randomplot, preddat)
se_diffs_randoms <- get_smooth_diffs(se.gam.randomplot.s, preddat)

se_preds <- predict_treat_effect2(plot_totals, 500, se.gam, exVars.d)
se_preds_noplot <- predict_treat_effect2(plot_totals, 500,se.gam.noplot, exVars.d)
se_preds_random <- predict_treat_effect2(plot_totals, 500,se.gam.randomplot, exVars.d)
se_preds_randoms <- predict_treat_effect2(plot_totals, 500,se.gam.randomplot.s, exVars.d)

ggplot(se_preds, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot smooth")


ggplot(se_diffs, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot smooth")

se_preds_fitted <- gratia::add_fitted(plot_totals, se.gam)

ggplot(se_preds_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()

```

The predictions here are misleading. Absent the plot-level smoothers, you end up with Ce exceeding EE ca. 2008 (for example). I believe this is because this model is able to use the plot smooths to correct for these weird offsets and get back to accurate fitted values. However, I worry this also makes the **difference** smooths also potentially suspect. For example, for comparing EE to the control, we cross 0 from 2005-2010. The **fitted values** are different, but the treatment effect smooths are crossing.

```{r, fig.dim = c(4,3)}
ggplot(filter(se_diffs, censusdate > "2005-01-01", censusdate < "2010-01-01", oTreatment == "EE"), aes(censusdate, diff, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot smooth")


ggplot(filter(se_preds, censusdate > "2005-01-01", censusdate < "2010-01-01", oTreatment != "CE"), aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot smooth")

ggplot(filter(plot_totals,censusdate > "2005-01-01", censusdate < "2010-01-01", oTreatment != "CE"), aes(censusdate, smgran_e, group = oPlot, color = oTreatment)) +
  geom_line()


ggplot(filter(se_preds_fitted,censusdate > "2005-01-01", censusdate < "2010-01-01", oTreatment != "CE"), aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()
```

More so, for this same time period, CE and CC ostensibly DO NOT OVERLAP. However the **fitted values** with plot do, and so do the real data. More than EE and CC do! 

```{r, fig.dim = c(4,3)}
ggplot(filter(se_diffs, censusdate > "2005-01-01", censusdate < "2010-01-01", oTreatment == "CE"), aes(censusdate, diff, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot smooth") +
  scale_colour_viridis_d(end = .4)


ggplot(filter(se_preds, censusdate > "2005-01-01", censusdate < "2010-01-01", oTreatment != "EE"), aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot smooth")+
  scale_colour_viridis_d(end = .4)

ggplot(filter(plot_totals,censusdate > "2005-01-01", censusdate < "2010-01-01", oTreatment != "EE"), aes(censusdate, smgran_e, group = oPlot, color = oTreatment)) +
  geom_line()+
  scale_colour_viridis_d(end = .4)


ggplot(filter(se_preds_fitted,censusdate > "2005-01-01", censusdate < "2010-01-01", oTreatment != "EE"), aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()+
  scale_colour_viridis_d(end = .4)
```

```{r}

ggplot(se_preds_random, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot random int")

ggplot(se_diffs_random, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random int")

se_preds_random_fitted <- gratia::add_fitted(plot_totals, se.gam.randomplot)

ggplot(se_preds_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()

```

This is a random-intercept plot model, and it does not have the same weird behaviors.




```{r}

ggplot(se_preds_randoms, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
 geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .25)+
  ggtitle("Plot random smooth")

se_preds_randoms_fitted <- gratia::add_fitted(plot_totals, se.gam.randomplot.s)

ggplot(se_preds_randoms_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()


ggplot(se_diffs_randoms, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random smooth")
```

Again, I'm not sure about this because it sure **looks** like the plot is being used to refine a very **vague** prediction. But at least it's not getting things dramatically wrong, like the rank order. (It really stands out having CE >> EE, as in the one with plot as a smooth)

```{r}

ggplot(se_preds_noplot, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5)+
  ggtitle("No plot")

ggplot(se_diffs_noplot, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("No plot")

```

Not including plot at all is pretty anti-conservative w.r.t. finding differences >0. 


```{r, fig.dim = c(4,3)}

ggplot(se_diffs_noplot, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("No plot")

ggplot(se_diffs_random, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random int")



```

Having the random plot effect, compared to no plot effect, gives a similar story but quite different CIs. 

- EE exceeds CC by quite a bit in the 1980s-1990s, declining to zero or near zero by the 2010s. With no plot, the smooths never actually cross; with plot, they cross about 2010.
- CE tracks CC more closely at first, but from 1995-around 2010 and from 2015 onwards, CE also exceeds CC.  

### Total E


```{r}

load(here::here("te_gams.RData"))

te_AICs <- lapply(list(te.gam, te.gam.noplot, te.gam.plotint, te.gam.randomplot, te.gam.randomplot.s), FUN = AIC)

exVars.d <- c('oPlot', paste0('s(numericdate):oPlot', unique(plot_totals$oPlot)))
preddat <- predict_treat_effect2(plot_totals, 500, te.gam, exVars.d) %>%
  select(censusdate, oPlot, oTreatment, numericdate)

te_diffs <- get_smooth_diffs(te.gam, preddat)
te_diffs_noplot <- get_smooth_diffs(te.gam.noplot, preddat)
te_diffs_random <- get_smooth_diffs(te.gam.randomplot, preddat)
te_diffs_randoms <- get_smooth_diffs(te.gam.randomplot.s, preddat)

te_preds <- predict_treat_effect2(plot_totals, 500, te.gam, exVars.d)
te_preds_noplot <- predict_treat_effect2(plot_totals, 500,te.gam.noplot, exVars.d)
te_preds_random <- predict_treat_effect2(plot_totals, 500,te.gam.randomplot, exVars.d)
te_preds_randoms <- predict_treat_effect2(plot_totals, 500,te.gam.randomplot.s, exVars.d)

ggplot(te_preds, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot smooth")


ggplot(te_diffs, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot smooth")

te_preds_fitted <- gratia::add_fitted(plot_totals, te.gam)

ggplot(te_preds_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()

```

Whoa, here it is again. The **actual** fits for CE don't wildly exceed CC and EE at any point, but absent the plot smooths they do. 


```{r}

ggplot(te_preds_random, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot random int")

ggplot(te_diffs_random, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random int")

te_preds_random_fitted <- gratia::add_fitted(plot_totals, te.gam.randomplot)

ggplot(te_preds_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()

```


```{r}

ggplot(te_preds_randoms, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
 geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .25)+
  ggtitle("Plot random smooth")

te_preds_randoms_fitted <- gratia::add_fitted(plot_totals, te.gam.randomplot.s)

ggplot(te_preds_randoms_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()


ggplot(te_diffs_randoms, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random smooth")
```

Again, the random smooths are weird...


```{r}

ggplot(te_preds_noplot, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5)+
  ggtitle("No plot")

ggplot(te_diffs_noplot, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("No plot")

```

No plot is anticonservative.

```{r, fig.dim = c(4,3)}

ggplot(te_diffs_noplot, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("No plot")

ggplot(te_diffs_random, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random int")


```

Again, having no plot vs. having the random plot effect tells a similar trend but with narrower/broader CIs.

- EE is well below the controls except for brief moments in the 2000s.
- CE is closer to the controls until 2015, when it jumps down to meet or nearly meet EE.

### Tinygran E


```{r}

load(here::here("tge_gams.RData"))

tge_AICs <- lapply(list(tge.gam, tge.gam.noplot, tge.gam.plotint, tge.gam.randomplot, tge.gam.randomplot.s), FUN = AIC)

exVars.d <- c('oPlot', paste0('s(numericdate):oPlot', unique(plot_totals$oPlot)))
preddat <- predict_treat_effect2(plot_totals, 500, tge.gam, exVars.d) %>%
  select(censusdate, oPlot, oTreatment, numericdate)

tge_diffs <- get_smooth_diffs(tge.gam, preddat)
tge_diffs_noplot <- get_smooth_diffs(tge.gam.noplot, preddat)
tge_diffs_random <- get_smooth_diffs(tge.gam.randomplot, preddat)
tge_diffs_randoms <- get_smooth_diffs(tge.gam.randomplot.s, preddat)

tge_preds <- predict_treat_effect2(plot_totals, 500, tge.gam, exVars.d)
tge_preds_noplot <- predict_treat_effect2(plot_totals, 500,tge.gam.noplot, exVars.d)
tge_preds_random <- predict_treat_effect2(plot_totals, 500,tge.gam.randomplot, exVars.d)
tge_preds_randoms <- predict_treat_effect2(plot_totals, 500,tge.gam.randomplot.s, exVars.d)

ggplot(tge_preds, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot smooth")


ggplot(tge_diffs, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot smooth")

tge_preds_fitted <- gratia::add_fitted(plot_totals, tge.gam)

ggplot(tge_preds_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()

```

I am absolutely not convinced that what we shoudl take from these data is that EE is **below** controls from 2007-2017!



```{r, fig.dim = c(4,3)}
ggplot(filter(tge_diffs, censusdate > "2005-01-01", censusdate < "2017-01-01", oTreatment == "EE"), aes(censusdate, diff, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot smooth")

ggplot(filter(tge_preds, censusdate > "2005-01-01", censusdate < "2017-01-01", oTreatment != "CE"), aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot smooth")

ggplot(filter(plot_totals,censusdate > "2005-01-01", censusdate < "2017-01-01", oTreatment != "CE"), aes(censusdate, tinygran_e, group = oPlot, color = oTreatment)) +
  geom_line()


ggplot(filter(tge_preds_fitted,censusdate > "2005-01-01", censusdate < "2017-01-01", oTreatment != "CE"), aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()
```


```{r}

ggplot(tge_preds_random, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5) +
  ggtitle("Plot random int")

ggplot(tge_diffs_random, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random int")

tge_preds_random_fitted <- gratia::add_fitted(plot_totals, tge.gam.randomplot)

ggplot(tge_preds_random_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()

```

I have no...objections here.


```{r}

ggplot(tge_preds_randoms, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
 geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .25)+
  ggtitle("Plot random smooth")

tge_preds_randoms_fitted <- gratia::add_fitted(plot_totals, tge.gam.randomplot.s)

ggplot(tge_preds_randoms_fitted, aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()


ggplot(tge_diffs_randoms, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random smooth")
```

```{r, fig.dim = c(4,3)}

ggplot(filter(tge_preds_randoms_fitted,censusdate > "2015-01-01"), aes(censusdate, .value, color = oTreatment, group = oPlot)) +
  geom_line()

ggplot(filter(plot_totals,censusdate > "2015-01-01"), aes(censusdate, tinygran_e, group = oPlot, color = oTreatment)) +
  geom_line()

```

These fits confuse me. The fitted values have CC strongly above CE and EE for this time period, but they're...not.

THIS IS BECAUSE THIS MODEL IS FIT WITH TOTAL E NOT TG E.


```{r}

ggplot(tge_preds_noplot, aes(censusdate, Fitted, color = oTreatment, fill = oTreatment)) +
  geom_line() +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = .5)+
  ggtitle("No plot")

ggplot(tge_diffs_noplot, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("No plot")

```


```{r, fig.dim= c(4,3)}


ggplot(tge_diffs_noplot, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("No plot")

ggplot(tge_diffs_random, aes(censusdate, diff, color = compare_trt, fill = compare_trt)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .5) +
  ggtitle("Plot random int")
```

These differ a bit. 

- EE way exceeds controls at the beginning, then **under** controls 2005-2010, then **over** again pretty quickly. With the random int the under/overs are not sig.
- CE tracks controls until either 2005-2010, when it too under performs, and then begins to exceed controls after 2015. 
