---
title: "Scaffolding analysis with plot switch data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
```

# Data

This document uses the data from the plots studied in Christensen et al 2019 ProcB. We will use some of the same control plots for the actual analysis, but both sets of treatment plots will be different. (Ch uses exclosures that get turned into controls and removals that get turned into controls. We will look at long term exclosures and controls that get turned into exclosures).

This gives us space to do some analytical development without wondering if we're seeing the actual effects. We have the same basic structures we will have in the actual data:

- Three treatment types - controls and two flavors of manipulation
- Changes to the manipulations occurring in 2015
- Probably similar sitewide dynamics in terms of fluctuations in sitewide abundance, major environmental events, etc. 
- Probably similar idiosyncracies in the data, e.g. variability between plots; autocorrelation; etc.

```{r}

rats <- read.csv(here::here("lore", "2020_redux", "christensen_plots.csv")) %>%
  mutate(plot = factor(plot),
         treatment = factor(treatment)) %>%
  filter(!is.na(species))

```

```{r}

rats_types_totals <- rats %>%
  #mutate(plot = as.ordered(plot),
 #        treatment = as.ordered(treatment)) %>%
  select(period, treatment, type, plot, energy) %>%
  group_by(period, treatment, type, plot) %>%
  summarize(nind = dplyr::n(),
            totale = sum(energy))  %>%
  ungroup() 


all_types <- expand.grid(period = unique(rats$period), plot = unique(rats$plot), type = unique(rats$type)) %>%
  left_join(distinct(select(rats, plot, treatment)))

rats_types_totals <- left_join(all_types, rats_types_totals)%>%
  mutate(nind = ifelse(is.na(nind), 0, nind),
         totale = ifelse(is.na(totale), 0, totale))
  

rats_totals <- rats %>% 
  mutate(plot = factor(plot),
         treatment = factor(treatment)) %>%
  select(period, treatment, plot, energy) %>%
  group_by(period, treatment, plot) %>%
  summarize(nind = dplyr::n(),
            totale = sum(energy)) %>%
  ungroup() 


all_totals <- expand.grid(period = unique(rats$period), plot = unique(rats$plot)) %>%
  left_join(distinct(select(rats, plot, treatment)))

rats_totals <- left_join(all_totals, rats_totals) %>%
  mutate(nind = ifelse(is.na(nind), 0, nind),
         totale = ifelse(is.na(totale), 0, totale))

major_changes <- c(118, 216, 356, 434) # beginning of Thibault data, Bailey's arrival, 2010 (changepoint), 2015 (treatments)

eras <- data.frame(
  period = unique(rats$period)
) %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "pre_th",
    ifelse(period <= 216, "pre_ba",
           ifelse(period <= 356, "pre_cpt",
                  ifelse(period <= 434, "pre_switch", NA))
  )))

```

Here is the raw data, crudely smoothed for visualization:

```{r}
ggplot(filter(rats_totals), aes(period, nind, color = treatment, group = plot)) +
 geom_smooth(se = F, size = 2) +  theme_bw() +
  scale_color_viridis_d(end = .8)  +ggtitle("Number of individuals")

# 
# ggplot(filter(rats_types_totals, type == "small_granivore"), aes(period, nind,  color = treatment, group = plot)) +
#  geom_smooth(se = F, size = 2) +  theme_bw() +
#   scale_color_viridis_d(end = .8)  +ggtitle("Number of individuals - small granivores")

ggplot(filter(rats_totals, treatment != "XC"), aes(period, totale, color = treatment, group = plot)) +
 geom_line() +  theme_bw() +
  scale_color_viridis_d(end = .8) + ggtitle("Energy use")



```

# 1. Does the energetic compensation persist over time?

## GAM

We fit a GAM to `energy ~ treatment + s(period) + s(period, by = treatment)` with a Tweedie link. We include all treatments and all time in this GAM, because we'll use it repeatedly. Treatment is ordered.

#### Details RMD is still working on

- Plot effect
- Ordered v. unordered factor

```{r}


energy <- filter(rats_totals) %>%
  mutate(oplot = ordered(plot),
         otreatment = as.ordered(treatment),
         type = "all") %>%
  as.data.frame()


library(mgcv)
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))

e.mod <- gam(totale ~  otreatment + s(period, k = 50) + s(period, by = otreatment, k = 50) + plot + s(period, k = 30, by = oplot), family = "tw", data  = energy)

summary(e.mod)

gam.check(e.mod)

e.pdat <- make_pdat(energy, include_plot = T, comparison_variable = "otreatment")

e.pred <- get_predicted_vals(e.mod, e.pdat)  %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era))


```

Here is the fitted values from that GAM:

```{r}

plot_fitted_pred(e.pred, comparison_variable = "otreatment") + ylab("Energy use") +
  geom_vline(xintercept = major_changes)
```

In specific "eras":

```{r}

plot_fitted_pred(filter(e.pred, era != "a_pre_th"), comparison_variable = "otreatment") + ylab("Energy use") +
  facet_wrap(vars(era), scales = "free_x", nrow = 1)

```