---
title: "PB"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.dim =c(8, 4))
library(soar)
library(ggplot2)
library(dplyr)

theme_set(theme_bw())
```

```{r}
plot_ps <- get_plot_totals(use_pre_switch = F)

treat_ps <- get_treatment_means(use_pre_switch = F) 
```

```{r}

ggplot(filter(plot_ps), aes(censusdate, pb_e_ma,group = plot, color = oplottype)) +
  geom_line() +
  facet_wrap(vars(oplottype)) +
  ggtitle("PB - all time")

ggplot(filter(plot_ps, as.numeric(oera) > 2), aes(censusdate, pb_e_ma,group = plot, color = oplottype)) +
  geom_line() +
  facet_wrap(vars(oplottype)) +
  ggtitle("PB - since 2010ish")



ggplot(filter(plot_ps), aes(censusdate, pb_e_ma/total_e_ma,group = plot, color = oplottype)) +
  geom_line() +
  facet_wrap(vars(oplottype)) +
  ggtitle("PB as percent of plot energy - all time") +
  geom_line(data = filter(treat_ps), aes(censusdate, pb_e_ma/total_e_ma, color = oplottype), inherit.aes = F, size = 2)

ggplot(filter(plot_ps, as.numeric(oera) > 2), aes(censusdate, pb_e_ma / total_e_ma,group = plot, color = oplottype)) +
  geom_line() +
  facet_wrap(vars(oplottype)) +
  ggtitle("PB as percent of plot energy - since 2010ish") +
  geom_line(data = filter(treat_ps, as.numeric(oera) > 2), aes(censusdate, pb_e_ma/total_e_ma, color = oplottype), inherit.aes = F, size = 2)
```

### models

```{r}

library(mgcv)
library(gratia)
library(emmeans)

plot_ps_pb <- plot_ps %>%
  mutate(pb_prop = pb_e / total_e) %>%
  filter(as.numeric(oera) > 1) %>%
  mutate(pb_prop = ifelse(is.nan(pb_prop), 0, pb_prop),
         row = dplyr::row_number())

qbin_int_gam <- gam(pb_prop ~ oera * oplottype + s(fplot, bs = "re"), family = quasibinomial(), data= plot_ps_pb)

qbin_int_gam_fit <- plot_ps_pb %>%
  add_fitted(qbin_int_gam, exclude = "s(fplot)")

ggplot(qbin_int_gam_fit, aes(censusdate, .value, color = oplottype)) +
  geom_line()

pairs(emmeans(qbin_int_gam, specs = ~ oera | oplottype))
plot(regrid(emmeans(qbin_int_gam, specs = ~ oera | oplottype)))
```

GLM:

```{r}
qbin_glm <- glm(pb_prop ~ oera * oplottype, family = quasibinomial(), data= plot_ps_pb)

summary(qbin_glm)
```


The above GLM struggles, I think because the controls in d are literally all 0. One could either remove just that time period for controls, or remove controls wholesale.


```{r}

qbin_glm_noctrl <- glm(pb_prop ~ oera * oplottype, family = quasibinomial(), data= filter(plot_ps_pb, oplottype != "CC"))

qbin_glm_nocd <- glm(pb_prop ~ oera * oplottype, family = quasibinomial(), data= filter(plot_ps_pb, paste0(oplottype, oera) != "CCd_post-switch"))

plot(pairs(emmeans(qbin_glm_noctrl, specs = ~ oera | oplottype)))
plot(pairs(emmeans(qbin_glm_nocd, specs = ~ oera | oplottype)))


```

It literally doesn't matter which you do, so I'm gonna go with no controls at all because the piecewise removal creeps me out. 


```{r}

ilink <- qbin_glm$family$linkinv

qbin_glm_se <- predict(qbin_glm_noctrl, type = "link", se.fit = T, newdata = filter(plot_ps_pb, oplottype != "CC")) %>%
  as.data.frame() %>%
  mutate(est = ilink(fit),
         lower = ilink(fit - 2*se.fit),
         upper = ilink(fit + 2*se.fit),
         period = filter(plot_ps_pb, oplottype != "CC")$period,
         oplottype = filter(plot_ps_pb, oplottype != "CC")$oplottype)


ggplot(qbin_glm_se, aes(period, est, color = oplottype, fill = oplottype)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax= upper), alpha = .3)


qbin_glm_se <- qbin_glm_se %>%
  right_join(select(plot_ps_pb, oera, oplottype, period))

qbin_glm_se %>%
  select(oera, oplottype, est, lower, upper) %>%
  distinct() %>%
  arrange(oplottype, oera)

treat_ps %>%
  group_by(oera, oplottype) %>%
  summarize(mean_pb_prop = mean(pb_e / total_e, na.rm = T)) %>%
  ungroup() %>%
  arrange(oplottype, oera)
```

KKKK, the above code is a mess, but things I think I'm learning...

- PB has declined sitewide, raw and as a proportion of total energy use. 
- It is now totally absent on control plots.
- There are a lot of modeling options and none of them are perfect. Some notes on the considerations:
    -~ The GAMs fit with a smooth for period and a random effect  for plot were doing a really bad job; they'd fit control higher than any other treatment even though that's just Wrong. This might be fixable by increasing the k for the period smooth, but I haven't gone deep in there because models with high k take a while to run and this isn't exactly what I'm trying to learn right now.~ I think this is because I wasn't including a parametric term for the factor.
    - I've been working with both betar() and quasibinomial() as GAM/GLM families. I think they give essentially the same results in this application, but haven't tested extensively.

What I really wanted to know was,

how much energy is PB using *now* on exclosures vs. when it was more dominant? And that looks like, an estimate now of about 14-18% post-2010 compared to about 60% pre-2010. If you don't take into account plot variation, and work just with the treatment means, that's about 24% post-2010. Compared to, essentially none on control plots. (Although note, oddly enough, it was more abundant on C --> E plots prior to the switch than on C --> C plots. I see this a bunch and I think it means we can't use those plots to get at priority effects.)




Since I've been working with GLS (allows for autocorr) in the ratios...

```{r}

library(nlme)

treat_ps <- treat_ps %>%
  mutate(pb_prop = pb_e / total_e) %>%
  group_by_all() %>%
  mutate(pb_prop = ifelse(is.na(pb_prop), 0, pb_prop)) %>%
  ungroup()

pb_treat <- filter(treat_ps, as.numeric(oera) != 1) %>%
  filter(oplottype != "CC")

pb_gls <- gls(pb_prop ~ plot_type * era, correlation = corAR1(form = ~ period | plot_type), data = pb_treat)

plot(emmeans(pb_gls, specs = ~ era | plot_type))

plot(pairs(emmeans(pb_gls, specs = ~ era | plot_type)))

```

The GLS makes some negative predictions for near zero values. It pushes the contrast for c-d for EC to not sig.  



The directly comparable model would be a GLM on treatment means. Again, including controls in d makes the CIs huge.

```{r}


pb_treat <- filter(treat_ps, as.numeric(oera) != 1) %>%
  filter(paste0(oplottype, oera) != "CCd_post-switch")  %>%
  group_by(oplottype) %>%
  mutate(pb_prop_ma = maopts(pb_prop)) %>%
  ungroup()

glm_treat <- glm(pb_prop ~ oera * oplottype, family = quasibinomial(), data= pb_treat)

plot(pairs(emmeans(glm_treat, specs = ~ oera | oplottype)))
plot(regrid(emmeans(glm_treat, specs = ~ oera | oplottype)))

ilink <- glm_treat$family$linkinv

glm_treat_se <- predict(glm_treat, type = "link", se.fit = T, newdata = pb_treat) %>%
  as.data.frame() %>%
  mutate(est = ilink(fit),
         lower = ilink(fit - 2*se.fit),
         upper = ilink(fit + 2*se.fit),
         period = filter(pb_treat)$period,
         oplottype = filter(pb_treat)$oplottype)


ggplot(glm_treat_se, aes(period, est, color = oplottype, fill = oplottype)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax= upper), alpha = .3)


glm_treat_se <- glm_treat_se %>%
  right_join(select(pb_treat, oera, oplottype, period, censusdate, pb_prop_ma))


ggplot(glm_treat_se, aes(censusdate, est, color = oplottype, fill = oplottype)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax= upper), alpha = .3) +
  geom_line(aes(y = pb_prop_ma)) +
  facet_grid(cols = vars(oera), space = "free_x", scales = "free_x") +
  ylab("PB % of energy use") +
  ggtitle("PB ratio over time") +
  theme(legend.position = "top")


```
