---
title: "Appendix S1 - Full analytical methods and model results"
subtitle: "Supplemental information for Diaz and Ernest, “Maintenance of community function through compensation breaks down over time in a desert rodent community”. In review at Ecology."
output:
  # html_document:
  #   toc: yes
  #   df_print: paged
  word_document:
    df_print: kable
    reference_docx: default_gdoc.docx
    toc: yes
  # # pdf_document:
  #   toc: yes
  #   df_print: kable
---

```{r setup, include=T, echo = F, results = F, warning = F, message = F}
knitr::opts_chunk$set(echo = F,  warning = F, message = F)

library(soar) # install using remotes::install_github("diazrenata/soar")
library(ggplot2)
library(dplyr)
library(multipanelfigure)
library(nlme)
library(emmeans)

theme_set(theme_bw())

```

\newpage

<!-- # Data -->

<!-- ## Functions -->

<!-- Data are accessed and processed using functions stored in https://github.com/diazrenata/soar, archived on Zenodo at https://doi.org/10.5281/zenodo.5539880. Install these functions either by running: -->


```{r, eval = F, echo = F}
remotes::install_github("diazrenata/soar")
```

<!-- or by downloading the Zenodo archive and installing the package manually.  -->

<!-- ## Data access -->

<!-- Data can be downloaded directly from the Portal data repository: -->

```{r, echo = F, eval = F}

plotl <- get_plot_totals()

plot_types <- list_plot_types() %>% filter(plot_type == "EE")
```

<!-- For speed and offline access, data files are also included in this repository in the `data` directory: -->

```{r, echo =F}

plotl <- read.csv(here::here("data", "plotl.csv"), stringsAsFactors = T)
plot_types <- read.csv(here::here("data", "plot_types.csv"), stringsAsFactors = T)

```

<!-- ## Balancing exclosure and control plots -->

<!-- Because there are 5 exclosure plots and 4 control plots in these data, we remove 1 exclosure plot to achieve a balanced design. From the 5 possible exclosures to remove, we randomly select 1 using the seed 1977 (the year the Portal Project was initiated).  -->

```{r, echo = F}

plot_types <- plot_types  %>% 
  filter(plot_type == "EE")

set.seed(1977) 
remove_plot <- sample(plot_types$plot, 1, F) # results in removing plot 19

plotl <- plotl %>%
  filter(plot != remove_plot)
```

<!-- ## Treatment-level means and quantities of interest -->

<!-- In order to calculate compensation and the total energy ratio, it is necessary to take the treatment-level mean total energy use and energy use by kangaroo rats and small granivores on control plots. For consistency in the main analysis, we take treatment-level means for all quantities. -->

<!-- Because this necessarily elides some degree of variability between plots with treatment types, we also conducted a provisional analysis incorporating between-plot variability for exclosure plots (but not for control plots), with qualitatively the same results (see appendix S4).  -->

<!-- To take treatment-level means: -->

```{r, echo = F}
# Treatment-level means:
treatl <- plots_to_treatment_means(plotl) 

# Format column types
treatl <- treatl %>%
  mutate(censusdate = as.Date(censusdate),
         oera = ordered(oera),
         oplottype = ordered(oplottype))
```


<!-- Calculate proportional energy use of *C. baileyi* on exclosure and control plots. The `pb_nozero` dataframe omits the first time period, because during that time *C. baileyi* was essentially absent at the site (and the large number of 0s for an entire treatment-by-factor level combination breaks statistical models). -->

```{r, echo = F}
pb <- get_pb(treatl) 

pb_nozero <- pb %>%
  filter(as.numeric(oera) > 1)

```

<!-- Calculate total energy ratio and compensation, comparing exclosure to control plots: -->

```{r, echo =F}
energy_ratio <- get_e_ratio(treatl)
compensation <- get_compensation(treatl)

```


<!-- Calculate kangaroo rat (Dipodomys) proportion of total energy use on control plots: -->

```{r, echo =F}

dipo_c_dat <- get_dipo_c(treatl)

```


<!-- ## Variable names for analyses -->

<!-- The variables used in these analyses, and their definitions. -->

<!-- - `period`: The monthly census period number for each census. Numeric. -->
<!-- - `censusdate`: The date of the monthly census. Date. -->
<!-- - `era`: The "time period", as described in the text. Character, one of `a_pre_pb` (first time period, before *C. baileyi* arrived at the site), `b_pre_reorg` (second time period, after *C. baileyi* established but before the most recent reorganization event), or `c_post_reorg` (third time period, after the last reorganization event). -->
<!-- - `oera`: `era` as an ordered factor, for modeling. Ordered factor. -->
<!-- - `plot_type`: The treatment, either `CC` for control or `EE` for exclosure. Character. -->
<!-- - `oplottype`: `plot_type` as an ordered factor, for modeling. Ordered factor. -->
<!-- - `total_e_rat`, `total_e_rat_ma` (specific to `energy_ratio`): The ratio of total energy use on exclosure plots relative to control plots, and the 6-month moving average. Numeric, unbounded. -->
<!-- - `smgran_comp`, `smgran_comp_ma` (specific to `compensation`): Energetic compensation by small granivores for kangaroo rat removal, and the 6-month moving average. Numeric, unbounded. -->
<!-- - `pb_prop`, `pb_prop_ma` (specific to `pb` and `pb_nozero`): The proportion of treatment-level energy use accounted for by *C. baileyi*, and the 6-month moving average. Numeric, proportion bounded 0-1. -->
<!-- - `dipo_prop`, `dipo_prop_ma` (specific to `dipo_c_dat`): The proportion of treatment-level energy use accounted for by all kangaroo rats, and the 6-month moving average. Numeric, proportion bounded 0-1. -->



\newpage

# Compensation

We fit a generalized least squares accounting for temporal autocorrelation between monthly censuses within each time period using a continuous autoregressive structure of order 1. We compared this model to models fit without the autocorrelation structure and without the time period term using AIC. The "full" model was the best-fitting model via AIC, and we used this model to calculate estimates and contrasts. 

```{r, echo = F}

comp_mean_gls <- gls(smgran_comp ~ oera,  correlation = corCAR1(form = ~ period), data = compensation)

comp_mean_gls_notime <- gls(smgran_comp ~ 1,  correlation = corCAR1(form = ~ period), data = compensation)

comp_mean_gls_noautoc <- gls(smgran_comp ~ oera, data = compensation)

comp_mean_null <- gls(smgran_comp ~ 1, data = compensation)
```


### Table S1. AIC scores for candidate models for compensation.

```{r, echo = F}


compensation_comparison<- data.frame(
  `Model specification` = c("intercept + timeperiod + autocorrelation",
                            "intercept + autocorrelation",
                            "intercept + timeperiod",
                            "intercept"),
  AIC = c(AIC(comp_mean_gls),
          AIC(comp_mean_gls_notime),
          AIC(comp_mean_gls_noautoc),
          AIC(comp_mean_null))
)


compensation_comparison

```


```{r, echo = F}

comp_mean_gls_emmeans <- emmeans(comp_mean_gls, specs = ~ oera)

```

### Table S1. Coefficients from GLS for compensation

```{r}
compensation_coef <- as.data.frame(summary(comp_mean_gls)$tTable)
compensation_coef
```


### Table S2. Estimates from GLS for compensation 

```{r}

compensation_estimates <- as.data.frame(comp_mean_gls_emmeans)
compensation_estimates

```

### Table S3. Contrasts from GLS for compensation 

```{r}
compensation_contrasts <- as.data.frame(pairs(comp_mean_gls_emmeans)) %>%
  mutate(p.value = round(p.value, digits = 4))
compensation_contrasts

```
\newpage

# Total energy use

As for compensation, we fit a generalized least squares accounting for temporal autocorrelation between monthly censuses within each time period using a continuous autoregressive structure of order 1. We compared this model to models fit without each of the terms, and found the full model had the best performance via AIC. We used this model for estimates and contrasts.

### Table S5. Model comparison via AIC for total energy ratio. 

```{r, echo = F}

totale_mean_gls <- gls(total_e_rat ~  oera,  correlation = corCAR1(form = ~ period), data = energy_ratio)


totale_mean_gls_notime <- gls(total_e_rat ~  1,  correlation = corCAR1(form = ~ period), data = energy_ratio)

totale_mean_gls_noautoc <- gls(total_e_rat ~ oera, data = energy_ratio)

totale_mean_null <- gls(total_e_rat ~ 1, data = energy_ratio)
```

```{r, echo =F}
totale_comparison<- data.frame(
  `Model specification` = c("intercept + timeperiod + autocorrelation",
                            "intercept + autocorrelation",
                            "intercept + timeperiod",
                            "intercept"),
  AIC = c(AIC(totale_mean_gls),
          AIC(totale_mean_gls_notime),
          AIC(totale_mean_gls_noautoc),
          AIC(totale_mean_null))
)


totale_comparison
```


<!-- We calculated estimates from this model using emmeans: -->

```{r, echo =F}
totale_mean_gls_emmeans <- emmeans(totale_mean_gls, specs = ~ oera)
```

### Table S4. Coefficients from GLS on total energy ratio

```{r}

te_coef <- as.data.frame(summary(totale_mean_gls)$tTable)

te_coef

```


### Table S5. Estimates from GLS on total energy ratio

```{r}

totale_estimates <- as.data.frame(totale_mean_gls_emmeans)
totale_estimates

```


### Table S6. Contrasts from GLS on total energy ratio

```{r}
totale_contrasts <- as.data.frame(pairs(totale_mean_gls_emmeans))%>%
  mutate(p.value = round(p.value, digits = 4))
totale_contrasts

```

\newpage

# Kangaroo rat proportional energy use

Proportional energy use is bounded 0-1 and cannot be fit with generalized least squares. We therefore use a binomial GLM with no temporal autocorrelation term. We compared a model fit with a timeperiod term to an intercept-only (null) model using AIC, and found the timeperiod term improved model fit. We used this model for estimates and contrasts.

```{r, echo = F}

dipo_glm <- glm(dipo_prop ~ oera, family = binomial, data= dipo_c_dat)

dipo_intercept_glm <- glm(dipo_prop ~ 1, family = binomial, data = dipo_c_dat)
```

### Table S9. Model comparison via AIC for Dipodomys proportional energy use.

```{r, echo =F}

dipo_comparison <- data.frame(
  `Model specification` = c("intercept + timeperiod",
                            "intercept"),
  AIC = c(AIC(dipo_glm),
          AIC(dipo_intercept_glm))
)


dipo_comparison
```


```{r, echo =F}
dipoemmeans <- regrid(emmeans(dipo_glm, specs = ~ oera))
```


### Table S7. Coefficients from GLM on Dipodomys energy use.

```{r}

dipo_coef <- as.data.frame(summary(dipo_glm)$coefficients)
dipo_coef

```


### Table S8. Estimates from GLM on Dipodomys energy use.

```{r}


dipoestimates <- as.data.frame(dipoemmeans)
dipoestimates

```


### Table S9. Contrasts from GLM on Dipodomys energy use.

```{r}
dipocontrasts <- as.data.frame(pairs(dipoemmeans))%>%
  mutate(p.value = round(p.value, digits = 4))
dipocontrasts

```

\newpage

# C. baileyi proportional energy use

## Model specification and selection


As for kangaroo rat energy use, we used a binomial GLM to fit *C. bailyei* proportional energy use. We investigated whether the dynamics *C. baileyi*'s proportional energy use differed between treatment types. We compared models incorporating separate slopes, separate intercepts, or no terms for treatment modulating the change in *C. baileyi* proportional energy use across time periods. We also tested a null (intercept-only) model of no change across time periods. We found that the best-fitting model incorporated effects for time period and for treatment, but no interaction between them. We therefore proceeded with this model. 

### Table S13; Model comparison via AIC for C. baileyi proportional energy use.


```{r, echo = F}

pb_glm_interaction <- glm(pb_prop ~ oera * oplottype, family = binomial, data= pb_nozero)
pb_glm_nointeraction <- glm(pb_prop ~ oera + oplottype, family = binomial, data= pb_nozero)
pb_glm_notreat <- glm(pb_prop ~ oera, family = binomial, data= pb_nozero)
pb_glm_notime <- glm(pb_prop ~ oplottype, family = binomial, data= pb_nozero)
pb_glm_null <- glm(pb_prop ~ 1, family = binomial, data= pb_nozero)

pb_comparison <- data.frame(
  `Model specification` = c("intercept + timeperiod + treatment + timeperiod:treatment",
                            "intercept + timeperiod + treatment",
                            "intercept + timeperiod",
                            "intercept + treatment",
                            "intercept"),
  AIC = c(AIC(pb_glm_interaction),
          AIC(pb_glm_nointeraction),
          AIC(pb_glm_notreat),
          AIC(pb_glm_notime),
          AIC(pb_glm_null))
)


pb_comparison

pb_emmeans <- regrid(emmeans(pb_glm_nointeraction, specs = ~ oera))
```

### Table S10. Coefficients from GLM on C. baileyi energy use

```{r}

pb_coef <- as.data.frame(summary(pb_glm_nointeraction)$coefficients)
pb_coef

```


### Table S11. Estimates from GLM on C. baileyi energy use

```{r}


pb_estimates <- as.data.frame(pb_emmeans)

pb_estimates

```

### Table S12. Contrasts from GLM on C. baileyi energy use.

```{r}
pb_contrasts <- as.data.frame(pairs(pb_emmeans)) %>%
  mutate(p.value = round(p.value, digits = 4))
pb_contrasts

```
