---
title: "pp, pb"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width =6, fig.height =3)
library(dplyr)
library(ggplot2)
```

# Data

```{r}

rats <- read.csv(here::here("lore", "2020_redux", "christensen_plots.csv")) %>%
  mutate(plot = factor(plot),
         treatment = factor(treatment)) %>%
  filter(!is.na(species))

```

```{r}

rats_totale<- rats %>%
  group_by(censusdate, period, treatment) %>%
  summarise(totale = sum(energy, na.rm = T)) %>%
  ungroup()

rats_nplots <- rats %>%
  select(plot, treatment) %>%
  distinct() %>%
  group_by(treatment) %>%
  summarize(nplots = length(unique(plot))) %>%
  ungroup()

rats_e_sp <- rats %>%
  group_by(censusdate, period, treatment, species) %>%
  summarize(spe = sum(energy, na.rm = T)) %>%
  ungroup() 


rats_all_e_sp <- expand.grid(period = unique(rats$period), species = c("PP", "PB"), treatment = unique(rats$treatment)) %>%
  left_join(distinct(select(rats, period, censusdate))) %>%
  left_join(rats_nplots)


rats_e_sp1 <- left_join(rats_all_e_sp, rats_e_sp) %>%
  left_join(rats_totale) %>%
  group_by_all() %>%
  mutate(spe = ifelse(is.na(spe), 0, spe),
         totale = ifelse(is.na(totale), 0, totale))  %>%
  ungroup() %>%
  mutate(prop_totale = spe/totale) %>%
  mutate(prop_totale = ifelse(is.nan(prop_totale), 0, prop_totale)) %>%
  mutate(mean_spe = (spe/nplots),
         mean_totale = (totale/nplots))
```

```{r}

ggplot(rats_e_sp1, aes(period, prop_totale, color = species)) + 
  geom_line() +
  facet_wrap(vars(treatment, species)) +
  theme_bw()

```

This is a mess. Trying to get ratio of (totale of PP on exclosure) to (totale on control) for PP, PB and exclosure, removal, and ratio of (totale on exclosure to totale on control). Then the ratio of species compensation to total compensation. Not sure. Trying to partition species contributions.

```{r}

totale_ctrl <- select(rats_e_sp1, period, treatment, censusdate, mean_totale) %>%
  filter(treatment == "CC") %>%
  select(-treatment) %>%
  rename(ctrl_mean_e = mean_totale) %>%
  distinct()

treatment_ratios <- select(rats_e_sp1, period, treatment, censusdate, mean_totale) %>%
  filter(treatment != "CC") %>%
  distinct() %>%
  left_join(totale_ctrl) %>%
  mutate(eratio = mean_totale / ctrl_mean_e)

species_ratios <- select(rats_e_sp1, period, treatment, species, mean_spe, mean_totale) %>%
  distinct() %>%
  filter(treatment != "CC") %>%
  left_join(totale_ctrl) %>%
  mutate(ctrl_ratio = mean_spe / ctrl_mean_e,
         trt_ratio = mean_spe / mean_totale,
         trt_to_ctrl = mean_totale/ctrl_mean_e) %>%
  mutate(trt_ratio = ifelse(mean_totale == 0, 0, trt_ratio),
         trt_to_ctrl = ifelse(ctrl_mean_e == 0, 0, trt_to_ctrl),
         ctrl_ratio = ifelse(ctrl_mean_e == 0, 0, ctrl_ratio)) %>%
  mutate(trt_ratio = ifelse(trt_ratio == 0, 0.00000001, trt_ratio),
         ctrl_ratio = ifelse(ctrl_ratio == 0, 0.00000001, ctrl_ratio)) %>%
  as.data.frame() %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era)) %>%
  group_by(treatment, era) %>%
  mutate(era_mean_trt_to_ctrl = mean(trt_to_ctrl)) %>%
  ungroup()

ggplot(filter(species_ratios, treatment == "EC"), aes(period, ctrl_ratio, color = species)) +
  geom_line() +
  theme_bw() +
  geom_line(aes(period, trt_to_ctrl, group =species), color = "black") +
  facet_wrap(vars(era), nrow = 1, scales = "free_x")



ggplot(filter(species_ratios, treatment == "EC"), aes(period, trt_ratio, color = species)) +
  geom_line() +
  theme_bw() +
  facet_wrap(vars(era), nrow = 1, scales = "free_x")


pp_glm <- glm(data = filter(species_ratios, species == "PP"),formula =  ctrl_ratio ~ era * treatment, family = Gamma)



library(lsmeans)

tt = lsmeans(pp_glm, specs = ~ era | treatment)
dd = pairs((tt))
ff = summary(dd)
dd_regrid = pairs(regrid(tt))
ff_regrid = summary(dd_regrid)
pp_glm_pred <- predict(pp_glm, se.fit = T)

ratio_pred <- species_ratios %>%
  filter(species == "PP")  %>%
  mutate(linkfit = pp_glm_pred$fit,
         linkupper = pp_glm_pred$se.fit + pp_glm_pred$fit,
         linklower = pp_glm_pred$fit -pp_glm_pred$se.fit) %>%
  mutate(invlinkfit = 1/linkfit,
         invlinklower = 1/linklower,
         invlinkupper = 1/linkupper)


ggplot(filter(ratio_pred), aes(period, ctrl_ratio, color = treatment, fill = treatment)) +
  geom_line() + 
  geom_line(aes(period, invlinkfit)) +
  geom_ribbon(aes(ymin = invlinklower, ymax = invlinkupper), alpha = .5) +
  facet_wrap(vars(era), scales = "free_x", nrow = 1) + 
  theme_bw() +
  scale_color_viridis_d(begin = .4, end = .8) +
    scale_fill_viridis_d(begin = .4, end = .8) +
  geom_hline(yintercept = 1, linetype = 3) +
  theme(legend.position = "top") +
  ylim(0, 2.5) +
  geom_line(aes(period, era_mean_trt_to_ctrl), color = "black") +
  ggtitle("PP")


pb_glm <- glm(data = filter(species_ratios, species == "PB"),formula =  ctrl_ratio ~ era * treatment, family = Gamma)



library(lsmeans)

tt = lsmeans(pb_glm, specs = ~ era | treatment)
dd = pairs((tt))
ff = summary(dd)
dd_regrid = pairs(regrid(tt))
ff_regrid = summary(dd_regrid)
pb_glm_pred <- predict(pb_glm, se.fit = T)

ratio_pred <- species_ratios %>%
  filter(species == "PB")  %>%
  mutate(linkfit = pb_glm_pred$fit,
         linkupper = pb_glm_pred$se.fit + pb_glm_pred$fit,
         linklower = pb_glm_pred$fit -pb_glm_pred$se.fit) %>%
  mutate(invlinkfit = 1/linkfit,
         invlinklower = 1/linklower,
         invlinkupper = 1/linkupper)


ggplot(filter(ratio_pred), aes(period, ctrl_ratio, color = treatment, fill = treatment)) +
  geom_line() + 
  geom_line(aes(period, invlinkfit)) +
  geom_ribbon(aes(ymin = invlinklower, ymax = invlinkupper), alpha = .5) +
  facet_wrap(vars(era), scales = "free_x", nrow = 1) + 
  theme_bw() +
  scale_color_viridis_d(begin = .4, end = .8) +
    scale_fill_viridis_d(begin = .4, end = .8) +
  geom_hline(yintercept = 1, linetype = 3) +
  theme(legend.position = "top") +
  ylim(0, 2.5)  +
  geom_line(aes(period, era_mean_trt_to_ctrl), color = "black") +
  ggtitle("PB")
```

I think this means there was a decline in PB, **offset somewhat by PP in a way that did not occur in b.**, but **not sufficient to offset the loss of PB**.