---
title: "Scaffolding analysis with plot switch data"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width =6, fig.height =3)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(dplyr)
library(ggplot2)
```

# Data

This document uses the data from the plots studied in Christensen et al 2019 ProcB. We will use some of the same control plots for the actual analysis, but both sets of treatment plots will be different. (Ch uses exclosures that get turned into controls and removals that get turned into controls. We will look at long term exclosures and controls that get turned into exclosures).

This gives us space to do some analytical development without wondering if we're seeing the actual effects. We have the same basic structures we will have in the actual data:

- Three treatment types - controls and two flavors of manipulation
- Changes to the manipulations occurring in 2015
- Probably similar sitewide dynamics in terms of fluctuations in sitewide abundance, major environmental events, etc. 
- Probably similar idiosyncracies in the data, e.g. variability between plots; autocorrelation; etc.

```{r}

rats <- read.csv(here::here("lore", "2020_redux", "christensen_plots.csv")) %>%
  mutate(plot = factor(plot),
         treatment = factor(treatment)) %>%
  filter(!is.na(species))

```

```{r}

rats_types_totals <- rats %>%
  #mutate(plot = as.ordered(plot),
 #        treatment = as.ordered(treatment)) %>%
  select(period, treatment, type, plot, energy) %>%
  group_by(period, treatment, type, plot) %>%
  summarize(nind = dplyr::n(),
            totale = sum(energy))  %>%
  ungroup() 


all_types <- expand.grid(period = unique(rats$period), plot = unique(rats$plot), type = unique(rats$type)) %>%
  left_join(distinct(select(rats, plot, treatment)))

rats_types_totals <- left_join(all_types, rats_types_totals)%>%
  mutate(nind = ifelse(is.na(nind), 0, nind),
         totale = ifelse(is.na(totale), 0, totale))
  

rats_totals <- rats %>% 
  mutate(plot = factor(plot),
         treatment = factor(treatment)) %>%
  select(period, treatment, plot, energy, censusdate) %>%
  group_by(period, treatment, plot, censusdate) %>%
  summarize(nind = dplyr::n(),
            totale = sum(energy)) %>%
  ungroup() 


all_totals <- expand.grid(period = unique(rats$period), plot = unique(rats$plot)) %>%
  left_join(distinct(select(rats, period, censusdate))) %>%
  left_join(distinct(select(rats, plot, treatment)))

rats_totals <- left_join(all_totals, rats_totals) %>%
  mutate(nind = ifelse(is.na(nind), 0, nind),
         totale = ifelse(is.na(totale), 0, totale))

major_changes <- c(118, 216, 356, 434) # beginning of Thibault data, Bailey's arrival, 2010 (changepoint), 2015 (treatments)

```

Here is the raw data (plotted as annual means per plot for visualization):

```{r}

rats_annual_totals <- rats_totals %>%
  mutate(censusdate = as.Date(censusdate)) %>%
  mutate(censusyear = as.numeric(format.Date(censusdate, "%Y"))) %>%
  group_by(censusyear, treatment, plot) %>%
  summarize(totale = sum(totale),
            nind = sum(nind),
            nsamples = length(unique(period))) %>%
  ungroup() %>%
  mutate(meane = totale/nsamples,
         meanind = nind/nsamples) %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    censusyear <= 1988, "a_pre_th",
    ifelse(censusyear <= 1996, "b_pre_ba",
           ifelse(censusyear <= 2010, "c_pre_cpt",
                  ifelse(censusyear <= 2015, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era))
# 
# ggplot(filter(rats_annual_totals), aes(censusyear, meanind, color = treatment, group = plot)) +
#  geom_line() +  theme_bw() +
#   scale_color_viridis_d(end = .8)  +ggtitle("Number of individuals") +
#   theme(legend.position = "top") +
#   facet_wrap(vars(era), scales = "free_x", nrow = 1)
#   

ggplot(filter(rats_annual_totals), aes(censusyear, meane, color = treatment, group = plot)) +
 geom_line() +  theme_bw() +
  scale_color_viridis_d(end = .8)  +ggtitle("Energy use")+
  theme(legend.position = "top")+
  facet_wrap(vars(era), scales = "free_x", nrow = 1)



```
## GAM

We will be working a lot with a GAM.

We fit a GAM to `energy ~ treatment + s(period) + s(period, by = treatment)` with a Tweedie link. We include all treatments and all time in this GAM, because we'll use it repeatedly. Treatment is ordered.

```{r}


energy <- filter(rats_totals) %>%
  mutate(oplot = ordered(plot),
         otreatment = as.ordered(treatment),
         type = "all") %>%
  as.data.frame() %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era))


library(mgcv)
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))

#e.mod <- gam(totale ~  otreatment + s(period, k = 100) + s(period, by = otreatment, k = 100), family = "tw", data  = energy)
e.mod <- gam(totale ~  otreatment + s(period, k = 20) + s(period, by = otreatment, k = 20), family = "tw", data  = energy)
# This k is too low but it takes forever to render with 100
#summary(e.mod)

#gam.check(e.mod)

e.pdat <- make_pdat(energy, include_plot = F, comparison_variable = "otreatment")

e.pred <- get_predicted_vals(e.mod, e.pdat)  %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era))


```

Here is the fitted values from that GAM:

```{r}

plot_fitted_pred(filter(e.pred), comparison_variable = "otreatment") + ylab("Energy use") +
  facet_wrap(vars(era), scales = "free_x", nrow = 1)

```


```{r}

cc_ec <- get_exclosure_diff(e.mod, e.pdat, comparison_variable = "otreatment") %>%
  left_join(select(e.pred, period, era))

# plot_exclosure_diff(filter(cc_ec)) +
#   facet_wrap(vars(era), scales = "free_x", nrow = 1)

e.newdiff <- add_exclosure_diff(e.pred, cc_ec)

plot_fitted_pred(filter(e.newdiff, otreatment != "XC"), comparison_variable = "otreatment") + ylab("Energy use") +
  facet_wrap(vars(era), scales = "free_x", nrow = 1) +
  scale_color_viridis_d(end = .4) +
  scale_fill_viridis_d(end = .4)

```


```{r}


ec_xc <- get_exclosure_diff(e.mod, e.pdat, reference_level = 2, comparison_level = 3, comparison_variable = "otreatment")

#plot_exclosure_diff(ec_xc)

e.origdiff <- add_exclosure_diff(e.pred, ec_xc)  %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era))

plot_fitted_pred(filter(e.origdiff, otreatment != "CC"), comparison_variable = "otreatment") + ylab("Energy use") +
  facet_wrap(vars(era), scales = "free_x", nrow = 1) +
  scale_color_viridis_d(begin = .4, end = .8) +
  scale_fill_viridis_d(begin = .4, end = .8)

```

## Comparing ratios

```{r}

energy_ratio <- energy %>%
  group_by(period, treatment, type, era, oera, otreatment) %>%
  summarize(totale = sum(totale),
            nplots = length(unique(plot))) %>%
  ungroup() %>%
  mutate(meane = totale/nplots) %>%
  tidyr::pivot_wider(id_cols = c(period, type, era, oera), names_from = treatment, values_from = meane) %>%
  mutate(ECr = EC/ CC,
         XCr = XC/CC) %>%
  dplyr::select(period, era, ECr, XCr) %>%
  tidyr::pivot_longer(cols = c(ECr, XCr), names_to = "ref", values_to = "ratio") %>%
  group_by(period,  ref) %>%
  mutate(mean_ratio = mean(ratio)) %>%
  ungroup() %>%
  mutate(ratio = ifelse(ratio == 0, 0.0000001, ratio)) %>%
  mutate(era = as.factor(era), ref = as.factor(ref))
# 
# ggplot(energy_ratio, aes(period, ratio, color = ref)) +
#   geom_line() + 
#   #geom_line(aes(period, mean_ratio)) + 
#   facet_wrap(vars(era), scales = "free_x", nrow = 1) + 
#   theme_bw() +
#   scale_color_viridis_d(begin = .4, end = .8) +
#   geom_hline(yintercept = 1) +
#   theme(legend.position = "top") +
#   ylim(0, 2.5) # this removes a couple points that stretch out the scale and make it hard to see. DO NOT DO THIS WITH ACTUAL DATA YOU WANT TO INTERPRET.
# 

```

```{r}
e_glm <- glm(ratio ~ era * ref, family = Gamma, data = energy_ratio)

library(lsmeans)

tt = lsmeans(e_glm, specs = ~ era | ref)
dd = pairs((tt))
ff = summary(dd)
dd_regrid = pairs(regrid(tt))
ff_regrid = summary(dd_regrid)
e_glm_pred <- predict(e_glm, se.fit = T)

ratio_pred <- energy_ratio %>%
  mutate(linkfit = e_glm_pred$fit,
         linkupper = e_glm_pred$se.fit + e_glm_pred$fit,
         linklower = e_glm_pred$fit -e_glm_pred$se.fit) %>%
  mutate(invlinkfit = 1/linkfit,
         invlinklower = 1/linklower,
         invlinkupper = 1/linkupper)
# 
# ggplot(filter(ratio_pred, ref == "ECr"), aes(period, ratio, color = ref, fill = ref)) +
#   geom_line() + 
#   geom_line(aes(period, invlinkfit)) +
#   geom_ribbon(aes(ymin = invlinklower, ymax = invlinkupper), alpha = .5) +
#   facet_wrap(vars(era), scales = "free_x", nrow = 1) + 
#   theme_bw() +
#   scale_color_viridis_d(begin = .4, end = .4) +
#     scale_fill_viridis_d(begin = .4, end = .4) +
#   geom_hline(yintercept = 1, linetype = 3) +
#   theme(legend.position = "top") +
#   ylim(0, 2.5) 



```

```{r}

ggplot(filter(ratio_pred), aes(period, ratio, color = ref, fill = ref)) +
  geom_line() + 
  geom_line(aes(period, invlinkfit)) +
  geom_ribbon(aes(ymin = invlinklower, ymax = invlinkupper), alpha = .5) +
  facet_wrap(vars(era), scales = "free_x", nrow = 1) + 
  theme_bw() +
  scale_color_viridis_d(begin = .4, end = .8) +
    scale_fill_viridis_d(begin = .4, end = .8) +
  geom_hline(yintercept = 1, linetype = 3) +
  theme(legend.position = "top") +
  ylim(0, 2.5) 
```

### Contrasts - in dev

From GLM with gamma link. RMD is exploring best practice for the modeling framework.

#### Era by treatment

```{r}


filter(ff, contrast %in% c("c_pre_cpt - d_pre_switch", "b_pre_ba - c_pre_cpt", "c_pre_cpt - (e_post-switch)", "d_pre_switch - (e_post-switch)"), ref == "ECr")

filter(ff_regrid, contrast %in% c("c_pre_cpt - d_pre_switch", "b_pre_ba - c_pre_cpt", "c_pre_cpt - (e_post-switch)", "d_pre_switch - (e_post-switch)"), ref == "ECr")


ratio_pred %>%
  select(era, ref, invlinkfit, invlinklower, invlinkupper) %>%
  distinct() %>%
  filter(ref == "ECr")

```

#### Treatment by era

```{r}
tt = lsmeans(e_glm, specs = ~ ref | era)
dd = pairs((tt))
ff = summary(dd)
dd_regrid = pairs(regrid(tt))
ff_regrid = summary(dd_regrid)

ff

ff_regrid

ratio_pred %>%
  select(era, ref, invlinkfit, invlinklower, invlinkupper) %>%
  distinct()

```

## PB

## PB abundance

```{r}

pb_totals <- rats %>% 
  mutate(plot = factor(plot),
         treatment = factor(treatment)) %>%
  filter(species == "PB") %>%
  select(period, treatment, plot, energy, censusdate) %>%
  group_by(period, treatment, plot, censusdate) %>%
  summarize(nind = dplyr::n(),
            totale = sum(energy)) %>%
  ungroup() 


all_totals <- expand.grid(period = unique(rats$period), plot = unique(rats$plot)) %>%
  left_join(distinct(select(rats, period, censusdate))) %>%
  left_join(distinct(select(rats, plot, treatment)))

pb_totals <- left_join(all_totals, pb_totals) %>%
  mutate(nind = ifelse(is.na(nind), 0, nind),
         totale = ifelse(is.na(totale), 0, totale))


pb_annual_totals <- pb_totals %>%
  mutate(censusdate = as.Date(censusdate)) %>%
  mutate(censusyear = as.numeric(format.Date(censusdate, "%Y"))) %>%
  group_by(censusyear, treatment, plot) %>%
  summarize(totale = sum(totale),
            nind = sum(nind),
            nsamples = length(unique(period))) %>%
  ungroup() %>%
  mutate(meane = totale/nsamples,
         meanind = nind/nsamples) %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    censusyear <= 1988, "a_pre_th",
    ifelse(censusyear <= 1996, "b_pre_ba",
           ifelse(censusyear <= 2010, "c_pre_cpt",
                  ifelse(censusyear <= 2015, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era))


ggplot(filter(pb_annual_totals), aes(censusyear, meanind, color = treatment, group = plot)) +
 geom_line() +  theme_bw() +
  scale_color_viridis_d(end = .8)  +ggtitle("PB Number of individuals") +
  theme(legend.position = "top") +
  facet_wrap(vars(era), scales = "free_x", nrow = 1)
  
```

```{r}


pb.nind <- filter(pb_totals) %>%
  mutate(oplot = ordered(plot),
         otreatment = as.ordered(treatment),
         type = "pb") %>%
  as.data.frame() %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era)) %>%
  filter(period > 216)


library(mgcv)
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))

#e.mod <- gam(totale ~  otreatment + s(period, k = 100) + s(period, by = otreatment, k = 100), family = "tw", data  = energy)
pb.mod <- gam(nind ~  otreatment + s(period, k = 20) + s(period, by = otreatment, k = 20), family = "poisson", data  = pb.nind)
# This k is too low but it takes forever to render with 100
#summary(e.mod)

#gam.check(e.mod)

pb.pdat <- make_pdat(pb.nind, include_plot = F, comparison_variable = "otreatment")

pb.pred <- get_predicted_vals(pb.mod, pb.pdat)  %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era))


plot_fitted_pred(filter(pb.pred), comparison_variable = "otreatment") + ylab("PB Abundance") +
  facet_wrap(vars(era), scales = "free_x", nrow = 1)

```

```{r}

pb_eras <- pb_totals  %>%
  group_by(period, treatment, censusdate) %>%
  summarize(nind = sum(nind)) %>%
  ungroup()

pb_all_eras <- expand.grid(treatment = unique(pb_totals$treatment),
                           period = unique(pb_totals$period)) 

pb_eras <- left_join(pb_all_eras, pb_eras) %>%
  mutate(nind = ifelse(is.na(nind), 0, nind)) %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era)) %>%
  filter(period > 216)


pb_glm <- glm(nind ~ era * treatment, family = poisson, data = pb_eras)

library(lsmeans)

tt = lsmeans(pb_glm, specs = ~ era | treatment)
dd = pairs((tt))
ff = summary(dd)
dd_regrid = pairs(regrid(tt))
ff_regrid = summary(dd_regrid)
pb_glm_pred <- predict(pb_glm, se.fit = T)

ilink <- pb_glm$family$linkinv

pb_mod_pred <- pb_eras %>%
  mutate(linkfit = pb_glm_pred$fit,
         linkupper = pb_glm_pred$se.fit + pb_glm_pred$fit,
         linklower = pb_glm_pred$fit -pb_glm_pred$se.fit) %>%
  mutate(invlinkfit = ilink(linkfit),
         invlinklower = ilink(linklower),
         invlinkupper = ilink(linkupper)) %>%
  group_by_all() %>%
  mutate(control_last = (treatment == "CC" && era == "e_post-switch")) %>%
  ungroup()

ggplot(filter(pb_mod_pred, !control_last), aes(period, nind, color = treatment, fill = treatment)) +
  geom_line() + 
  geom_line(aes(period, invlinkfit)) +
 geom_ribbon(aes(ymin = invlinklower, ymax = invlinkupper), alpha = .5) +
  facet_wrap(vars(era), scales = "free_x", nrow = 1) + 
  theme_bw() +
  scale_color_viridis_d(end = .8) +
    scale_fill_viridis_d(end = .8) +
  theme(legend.position = "top")

ff

ff_regrid


pb_mod_pred %>%
  select(era, treatment, invlinkfit, invlinklower, invlinkupper) %>%
 distinct()


```

#### pb with plot

```{r}


pb_totals_eras <- pb_totals  %>%
  mutate(nind = ifelse(is.na(nind), 0, nind)) %>%
  mutate(era = NA) %>%
  mutate(era = ifelse(
    period <= 118, "a_pre_th",
    ifelse(period <= 216, "b_pre_ba",
           ifelse(period <= 356, "c_pre_cpt",
                  ifelse(period <= 434, "d_pre_switch", "e_post-switch"))))) %>%
  mutate(oera = as.ordered(era)) %>%
  filter(period > 216)

pb_glm <- glm(nind ~ era * treatment + plot, family = poisson, data = pb_totals_eras)

library(lsmeans)

tt = lsmeans(pb_glm, specs = ~ era | treatment)
dd = pairs((tt))
ff = summary(dd)
dd_regrid = pairs(regrid(tt))
ff_regrid = summary(dd_regrid)
pb_glm_pred <- predict(pb_glm, se.fit = T)

ilink <- pb_glm$family$linkinv

pb_mod_pred <- pb_totals_eras %>%
  mutate(linkfit = pb_glm_pred$fit,
         linkupper = pb_glm_pred$se.fit + pb_glm_pred$fit,
         linklower = pb_glm_pred$fit -pb_glm_pred$se.fit) %>%
  mutate(invlinkfit = ilink(linkfit),
         invlinklower = ilink(linklower),
         invlinkupper = ilink(linkupper)) %>%
  group_by_all() %>%
  mutate(control_last = (treatment == "CC" && era == "e_post-switch")) %>%
  ungroup()

ggplot(filter(pb_mod_pred, !control_last), aes(period, nind, color = treatment, fill = treatment)) +
  geom_line() + 
  geom_line(aes(period, invlinkfit)) +
 geom_ribbon(aes(ymin = invlinklower, ymax = invlinkupper), alpha = .5) +
  facet_wrap(vars(era), scales = "free_x", nrow = 1) + 
  theme_bw() +
  scale_color_viridis_d(end = .8) +
    scale_fill_viridis_d(end = .8) +
  theme(legend.position = "top")

ff

ff_regrid


pb_mod_pred %>%
  select(era, treatment, plot, invlinkfit, invlinklower, invlinkupper) %>%
  group_by(era, treatment) %>%
  summarize(fitmean = mean(invlinkfit))


pb_mod_pred %>%
  select(era, treatment, plot, invlinkfit, invlinklower, invlinkupper) %>%
 distinct()

```