---
title: "MS figs"
output:
  github_document:
       df_print: kable
       toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.dim = c(6, 3))
library(soar)
library(ggplot2)
library(dplyr)
library(mgcv)
library(gratia)
theme_set(theme_bw())

library(nlme)
library(emmeans)
treatl <- get_treatment_means() %>%
  remove_switch()
both_scale <- scale_color_viridis_d(option = "cividis", begin = .1, end = .8)
both_fscale <- scale_fill_viridis_d(option = "cividis", begin = .1, end = .8)


pb <- get_pb(treatl)
pb_nozero <- pb %>%
  filter(as.numeric(oera) > 1)
energy_ratio <- get_e_ratio(treatl)
compensation <- get_compensation(treatl)

era_df <- make_era_df()


```

# Figures

## 1. Compensation and total energy use


Lines are 6-month moving averages. Horizontal lines + ribbons are means and SE or CL from GLM or GLS.

**Compensation** refers to compensatory gains in energy use by small granivores on exclosure plots relative to controls. Calculated as $\frac{SmgranExclosure - SmgranControl}{DipoControl}$. **Total energy** refers to the overall loss in energy use caused by kangaroo rat removal.

```{r}


comp_mean_gls <- gls(smgran_comp ~ era,  correlation = corCAR1(form = ~ period), data = compensation)

comp_mean_gls_emmeans <- emmeans(comp_mean_gls, specs = ~ era)

comp_mean_pred <-  as.data.frame(comp_mean_gls_emmeans) %>%
  right_join(compensation)

comp_plot <- ggplot(filter(comp_mean_pred, oplottype %in% c("CC", "EE")), aes(censusdate, emmean)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = .2) +
  geom_line(aes(y = smgran_comp_ma)) +
  ggtitle("Energetic compensation for kangaroo rat removal") +
  ylab("(Smgran_trt - Smgran_ctrl) / Dipo_ctrl") +
  ylim(-.05, 1) +
  geom_segment(data = era_df, aes(x = event_date, xend = event_date, y = -.05, yend = 1), linetype = 3)+
  xlab("") +
  scale_x_date(sec.axis = dup_axis(name = NULL, breaks = era_df$event_date, labels = era_df$event_name)) +
  theme(axis.title.y = element_text(size = 8))

#comp_plot
```


```{r}

totale_mean_gls <- gls(total_e_rat ~  era,  correlation = corCAR1(form = ~ period), data = energy_ratio)

totale_mean_gls_emmeans <- emmeans(totale_mean_gls, specs = ~ era)


totale_mean_pred <- as.data.frame(totale_mean_gls_emmeans) %>%
  right_join(energy_ratio)
  
totale_plot <- ggplot(filter(totale_mean_pred, oplottype %in% c("CC", "EE")), aes(censusdate, emmean)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = .2) +
  geom_line(aes(y = total_e_rat_ma)) +
  ggtitle("Total energy use on exclosures relative to controls") +
  ylab("Exclosure / control energy use") +
  theme(legend.position = "none") +
  ylim(-0, 1.2) +
  xlab("Date")  +
  geom_segment(data = era_df, aes(x = event_date, xend = event_date, y = 0, yend = 1.2), linetype = 3) +
  scale_x_date(sec.axis = dup_axis(name = NULL, breaks = era_df$event_date, labels = era_df$no_name))+
  theme(axis.title.y = element_text(size = 8))

#totale_plot

```

```{r, fig.dim = c(6,4)}

gridExtra::grid.arrange(grobs = list(comp_plot, totale_plot), ncol = 1)

```

## 2. Rodent community composition



```{r}


pb_glm_treat <- glm(pb_prop ~ oera * oplottype, family = quasibinomial(), data= pb_nozero)

pb_glm_treat_se <- est_glm_ilink(pb_glm_treat, pb_nozero) %>%
  full_join(select(pb, period, oplottype, pb_prop_ma, censusdate)) %>%
  mutate(Treatment = ifelse(oplottype == "CC", "Control", "Exclosure"),
         Rodents = "C. baileyi") %>%
  rename(rod_prop_ma = pb_prop_ma)


smgran_dat <- treatl %>%
  mutate(smgran_prop = smgran_e / total_e) %>%
  group_by(oplottype) %>%
  mutate(smgran_prop_ma = maopts(smgran_prop)) %>%
  ungroup() 

smgran_glm_treat <- glm(smgran_prop ~ oera * oplottype, family = quasibinomial(), data= smgran_dat)

smgran_glm_treat_se <- est_glm_ilink(smgran_glm_treat, smgran_dat) %>%
  left_join(select(smgran_dat, period, oplottype, smgran_prop_ma)) %>%
  mutate(Treatment = ifelse(oplottype == "CC", "Control", "Exclosure"),
         Rodents = "All small granivores\n(non-Dipodomys)") %>%
  rename(rod_prop_ma = smgran_prop_ma)

both_treat_se <- bind_rows(pb_glm_treat_se, smgran_glm_treat_se)

cplot <- ggplot(filter(both_treat_se, oplottype == "CC"), aes(censusdate, est, color = Rodents, fill = Rodents)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax= upper), alpha = .3) +
  geom_line(aes(y = rod_prop_ma)) +
  ylab("Proportion of total community energy use") +
  ggtitle("Controls") +
  theme(legend.position = c(.15, .6)) + both_scale + both_fscale   +
  geom_segment(data = era_df, aes(x = event_date, xend = event_date, y = 0, yend = 1), linetype = 3, inherit.aes = F) +
  scale_x_date(sec.axis = dup_axis(name = NULL, breaks = era_df$event_date, labels = era_df$event_name)) + xlab("") +
  theme(axis.title.y = element_text(size = 8), legend.title = element_blank(), legend.text = element_text(size =8))+ theme(axis.title.x = element_blank(), axis.title.y = element_blank(), title = element_text(size = 10), legend.background = element_rect(fill = NULL))

eplot <-  ggplot(filter(both_treat_se, oplottype == "EE"), aes(censusdate, est, color = Rodents, fill = Rodents)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax= upper), alpha = .3) +
  geom_line(aes(y = rod_prop_ma)) +
  ylab("Proportion of total community energy use") +
  ggtitle("Kangaroo rat exclosures") +
  theme(legend.position = "none") + both_scale + both_fscale   +
  geom_segment(data = era_df, aes(x = event_date, xend = event_date, y = 0, yend = 1), linetype = 3, inherit.aes = F) +
  scale_x_date(sec.axis = dup_axis(name = NULL, breaks = era_df$event_date, labels = era_df$no_name)) + xlab("Date") +
  theme(axis.title.y = element_text(size = 8)) +theme(axis.title.x = element_blank(), axis.title.y = element_blank(), title = element_text(size = 10))


```

```{r, fig.dim = c(6,4)}
library(grid)
gridExtra::grid.arrange(grobs = list(cplot, eplot), ncol = 1, top = textGrob("Rodent community composition"), bottom = textGrob("Date"), left = textGrob("Proportion of total community energy use", rot = 90))

```

# Model results

## 1. Compensation & total energy use

### Compensation

```{r}

summary(comp_mean_gls)
```


Estimates: 

```{r}

compensation_estimates <- as.data.frame(comp_mean_gls_emmeans)
compensation_estimates

```

Contrasts:

```{r}
compensation_contrasts <- as.data.frame(pairs(comp_mean_gls_emmeans))
compensation_contrasts

```

### Total energy use


```{r}

summary(totale_mean_gls)
```


Estimates: 

```{r}

totale_estimates <- as.data.frame(totale_mean_gls_emmeans)
totale_estimates

```

Contrasts:

```{r}
totale_contrasts <- as.data.frame(pairs(totale_mean_gls_emmeans))
totale_contrasts

```

## 2. Community composition

### All small granivores


```{r}

summary(smgran_glm_treat)
```


Estimates: 

Estimates from `emmeans` differ numerically (very slightly) from estimates obtained via `predict()` and back transformation. Below are estimates from `emmeans`, because those are what are used for contrasts.

```{r}


smgran_emmeans <- regrid(emmeans(smgran_glm_treat, specs = ~ oera | oplottype))

smgran_estimates <- as.data.frame(smgran_emmeans)

smgran_estimates

```

Estimates from `predict`:

```{r}

select(smgran_glm_treat_se, oplottype, oera, est, lower, upper) %>% distinct()

```


Contrasts:

```{r}
smgran_contrasts <- as.data.frame(pairs(smgran_emmeans))
smgran_contrasts

```


### C. baileyi


```{r}

summary(pb_glm_treat)
```


Estimates: 

```{r}

pb_emmeans <- regrid(emmeans(pb_glm_treat, specs = ~ oera | oplottype))

pb_estimates <- as.data.frame(pb_emmeans)

pb_estimates

```


Estimates from `predict`:

```{r}

select(pb_glm_treat_se, oplottype, oera, est, lower, upper) %>% distinct()

```

Contrasts:

```{r}
pb_contrasts <- as.data.frame(pairs(pb_emmeans))
pb_contrasts

```