---
title: "Treatments 1977-1991"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rats <- read.csv(here::here("lore", "1994_longterm", "1994_data_complete.csv"), stringsAsFactors = F)
rat_totals_1977 <- read.csv(here::here("lore", "1994_longterm", "1994_data_statevars_1977.csv"), stringsAsFactors = F)
rat_totals_1988 <- read.csv(here::here("lore", "1994_longterm", "1994_data_statevars_1988.csv"), stringsAsFactors = F)
library(dplyr)
library(ggplot2)

```


This is as best as RMD can reconstruct based on the text of Brown & Munger 1981 and Heske, Brown, and Mistry 1994, and the [PortalData plot treatments table](https://github.com/weecology/PortalData/blob/master/SiteandMethods/Portal_plot_treatments.csv).


```{r}

plots <- data.frame(plot = as.factor(c(1:24)),
                    treatment = NA)

plots$treatment[c(3,15,19,21)] <- "orig_exclosure"
plots$treatment[c(8,11,12,14)] <- "orig_control"
plots$treatment[c(6,13,18,20)] <- "second_exclosure"
plots$treatment[c(2,4,17,22)] <- "second_control"

plots <- plots %>%
  mutate(seventies = ifelse(treatment == "orig_exclosure", "exclosure", "control"),
         eighties = ifelse(grepl(treatment, pattern = "exclosure"), "exclosure", ifelse(is.na(treatment), NA, "control"))) %>%
  filter(!is.na(treatment)) %>%
  tidyr::pivot_longer(cols = c(seventies, eighties), names_to = "time_period") %>%
  mutate(time_period = ordered(time_period, levels = c("seventies", "eighties"))) %>%
  mutate(paper_period = ifelse(grepl(treatment, pattern = "orig"), "Brown and Munger 1981", "Heske et al 1994")) %>%
  mutate(paper_period = ordered(paper_period, levels = c("Brown and Munger 1981", "Heske et al 1994")))

timeperiods <- data.frame(year = c(1977:1991)) %>%
  mutate(time_period = ifelse(year < 1988, "seventies", "eighties"))

plots <- right_join(plots, timeperiods)

ggplot(plots, aes(year, plot, color = value)) +
  geom_line(size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(paper_period), ncol = 1)


ggplot(plots, aes(year, plot, color = value, linetype = paper_period)) +
  geom_line(size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) 
```


Looking ahead to 2015...


```{r}
plots_longer <- data.frame(plot = c(1:24),
                           initial = NA,
                           nineties = NA,
                           aughts = NA,
                           now = NA)

plots_longer$initial <- c("ctrl", "ctrl", "excl", "ctrl", "spec", "ctrl", "rmvl", "ctrl", "ctrl", "rmvl", "ctrl", "ctrl", "ctrl", "ctrl", "excl", "rmvl", "ctrl", "ctrl", "excl", "ctrl", "excl", "ctrl", "rmvl", "spec")

plots_longer$nineties <- c("spec", "ctrl", "excl", "ctrl", "rmvl", "excl", "rmvl", "ctrl", "spec", "rmvl", "ctrl", "ctrl", "excl", "ctrl", "excl", "rmvl", "ctrl", "excl", "excl", "excl", "excl", "ctrl", "rmvl", "rmvl")

plots_longer$aughts <- c("ctrl", "ctrl", "excl", "ctrl", "rmvl", "excl", "rmvl", "ctrl", "ctrl", "rmvl", "ctrl", "ctrl", "excl", "ctrl", "excl", "rmvl", "ctrl", "excl", "excl", "excl", "excl", "ctrl", "rmvl", "rmvl")


plots_longer$now <- c("rmvl", "excl", "excl", "ctrl", "ctrl", "ctrl", "ctrl", "excl", "rmvl", "rmvl", "ctrl", "rmvl", "ctrl", "ctrl", "excl", "rmvl", "ctrl", "ctrl", "excl", "excl", "excl", "excl", "rmvl", "ctrl")

longer_timeperiods <- data.frame(year = 1977:2020, timeperiod = NA)

longer_timeperiods <- longer_timeperiods %>%
  mutate(timeperiod = ifelse(
    year < 1988, "initial",
    ifelse(year < 2005, "nineties",
           ifelse(year < 2015, "aughts",
           "now")
  )
  )
  )

plots_longer <- plots_longer %>%
  tidyr::pivot_longer(-plot, names_to = "timeperiod", values_to = "treatment")

plots_longer_years <- left_join(longer_timeperiods, plots_longer) %>%
  mutate(plot = factor(plot))

ggplot(plots_longer_years, aes(year, plot, color = treatment)) +
  geom_line() +
  theme_bw() +
  scale_color_viridis_d(end = .8)

paper_treatments <- data.frame(plot = c(1:24), brown_treatment = NA, heske_treatment = NA, chr_treatment = NA, diaz_treatment = NA) %>%
  mutate(plot = as.factor(plot))

paper_treatments$heske_treatment[c(2, 4, 17, 22)] <- "ctrl"

paper_treatments$heske_treatment[c(6, 13, 18, 20)] <- "excl"

paper_treatments$brown_treatment[c(8, 11, 12, 14)] <- "ctrl"

paper_treatments$brown_treatment[c(3, 15, 19, 21)] <- "excl"

paper_treatments$chr_treatment[c(4, 11, 14, 17)] <- "ctrl"

paper_treatments$diaz_treatment[c(4, 11, 14, 17)] <- "ctrl"
paper_treatments$diaz_treatment[c(2, 8, 22)] <- "excl"
paper_treatments$diaz_treatment[c(3, 15, 19, 21)] <- "orig_excl"
paper_treatments$diaz_treatment[c(20)] <- "heske_excl"

plots_longer_years <- left_join(plots_longer_years, paper_treatments)


ggplot(plots_longer_years, aes(year, plot, color = treatment)) +
  geom_line(size = 2, alpha = .6) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  xlim(1976, 2022) +
  geom_point(data = filter(plots_longer_years, !is.na(brown_treatment)), aes(1981, plot, shape = brown_treatment), color = "black", size = 3) +
  geom_point(data = filter(plots_longer_years, !is.na(heske_treatment)), aes(1994, plot, shape = heske_treatment), color = "black", size = 3) +
  geom_point(data = filter(plots_longer_years, !is.na(chr_treatment)), aes(2019, plot, shape = chr_treatment), color = "black", size = 3)+
  geom_point(data = filter(plots_longer_years, !is.na(diaz_treatment)), aes(2021, plot, shape = diaz_treatment), color = "black", size = 3)
```

Given the structure here. I think the 2015 analysis comparisons would be 

* Comparing new exclosures to (old) controls
* Comparing old exclosures to (old) controls
* Comparing new exclosures to old exclosures
    * This is maybe the most interesting one. 
    * It's a little different from Heske because Heske has 2 sets of control plots and by 2015 we just have one.
    * Particularly interested in the **difference** between new excl-ctrl vs old excl-ctrl. In this instance this is going to be the same as comparing the two exclosures because the control plots are the same.
    * Do new exclosures match old exclosures, and does this happen rapidly or after a time lag?
        * Possible divergence in the plant communities over time on the plots
        * Also curious about contributions from PB
    * This comparison kind of encompasses what we learn from the first two. In particular, the magnitude of whatever response does or doesn't occur on the new exclosures is most meaningful *compared to* the magnitude of the difference we're seeing between old exclosures and controls. To the extent that that reflects a "baseline" capacity for the small granivore community at this point in time. Over or underperformance means the new exclosures are extra good/bad; matching means there isn't a strong habitat effect on net. 
* Comparing new exclosures, pre-treatment, to (old) controls
    * This is a check to confirm that, prior to the switch, the plots that end up being converted are not systematically different from the ones that stay controls.
    

Also, given the checkered history of the plots - the designations here are specifically **not** including various ant, plant, and seed treatments in the first few decades - I think it is **very important** to include the plot effect. 


I had some thought that maybe it doesn't make sense to include treatment as an **ordered** factor because the controls differ. That's the case for the Heske analysis but not for 2015, so for now I will stick with the ordered factor.