---
title: "Small granivores originals"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rats <- read.csv(here::here("lore", "1994_longterm", "1994_data_complete.csv"), stringsAsFactors = F)
rat_totals_1977 <- read.csv(here::here("lore", "1994_longterm", "1994_data_statevars_1977.csv"), stringsAsFactors = F)
rat_totals_1988 <- read.csv(here::here("lore", "1994_longterm", "1994_data_statevars_1988.csv"), stringsAsFactors = F)
library(dplyr)
library(ggplot2)

```

```{r}

rat_totals <- rat_totals_1977 %>%
  rename(trtmnt = trtmnt_1977) %>%
  mutate(time_period = "seventies") %>%
  bind_rows(mutate(rename(rat_totals_1988, trtmnt = trtmnt_1988), time_period = "eighties"))
```


```{r}
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))

sg_all<- filter(rat_totals, type == "small_granivore") %>%
  mutate(plot_type = paste(time_period, trtmnt, sep = "_")) %>%
  mutate(oplot_type = ordered(plot_type, levels = c("seventies_control", "seventies_exclosure", "eighties_control", "eighties_exclosure")))

ggplot(sg_all, aes(as.Date(censusdate), nind, color = oplot_type)) +
  geom_line(size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) + 
  theme(legend.position = "top")

library(mgcv)

sg.gam <- gam(nind ~ oplot_type + s(period, k  = 30) + s(period, by = oplot_type, k = 30), family = "poisson", data = sg_all)
summary(sg.gam)

gam.check(sg.gam)

sg.pdat <- make_pdat(sg_all, comparison_variable = "oplot_type")
sg.pred <- get_predicted_vals(sg.gam, sg.pdat)


#sg.pred <- add_exclosure_diff(sg.pred, sg.diff)

print(plot_link_pred(sg.pred, comparison_variable = "oplot_type"))

print(plot_fitted_pred(sg.pred, comparison_variable = "oplot_type"))


seventies_pred <- filter(sg.pred, oplot_type %in% c("seventies_control", "seventies_exclosure"))
print(plot_fitted_pred(seventies_pred, comparison_variable = "oplot_type"))

seventies_diff <- get_exclosure_diff(sg.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 1, comparison_level = 2)

print(plot_exclosure_diff(seventies_diff))

print(plot_fitted_pred(add_exclosure_diff(seventies_pred, seventies_diff), comparison_variable = "oplot_type"))

seventies_diff %>%
  filter(diff_overlaps_zero) %>%
  filter(period == max(period)) %>%
  distinct()

eighties_pred <- filter(sg.pred, oplot_type %in% c("eighties_control", "eighties_exclosure"))

print(plot_fitted_pred(eighties_pred, comparison_variable = "oplot_type"))

eighties_diff <- get_exclosure_diff(sg.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 3, comparison_level = 4)

print(plot_exclosure_diff(eighties_diff))

eighties_pred <- add_exclosure_diff(eighties_pred, eighties_diff)
print(plot_fitted_pred(eighties_pred, comparison_variable = "oplot_type"))

max(eighties_pred$period [ which(eighties_pred$diff_overlaps_zero)])

filter(rat_totals, period > 116, period < 122) %>%
  select(period, censusdate) %>%
  distinct()


controls_pred <- filter(sg.pred, oplot_type %in% c("eighties_control", "seventies_control"))
print(plot_fitted_pred(controls_pred, comparison_variable = "oplot_type"))

controls_diff <- get_exclosure_diff(sg.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 1, comparison_level = 3)
print(plot_exclosure_diff(controls_diff))

controls_pred <- add_exclosure_diff(controls_pred, controls_diff)

print(plot_fitted_pred(controls_pred, comparison_variable = "oplot_type"))

exclosures_pred <- filter(sg.pred, oplot_type %in% c("eighties_exclosure", "seventies_exclosure"))
print(plot_fitted_pred(exclosures_pred, comparison_variable = "oplot_type"))

exclosures_diff <- get_exclosure_diff(sg.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 2, comparison_level = 4)
print(plot_exclosure_diff(exclosures_diff))

exclosures_pred <- add_exclosure_diff(exclosures_pred, exclosures_diff)
print(plot_fitted_pred(exclosures_pred, comparison_variable = "oplot_type"))

exclosures_pred %>%
  filter(period > 100) %>%
  filter(diff_overlaps_zero) %>%
  select(period) %>%
  distinct() %>%
  filter(period == min(period))

filter(rat_totals, period > 116, period < 122) %>%
  select(period, censusdate) %>%
  distinct()


# 
# sg.diff <- get_exclosure_diff(sg.gam, sg.pdat,comparison_variable = "trtmnt")
# 
# print(plot_exclosure_diff(sg.diff) +
#         geom_vline(xintercept = 35))
# 
# 
# min(sg.diff$period[ which(!sg.diff$diff_overlaps_zero)])

```


In this model (which has no "plot" effect):

- The two sets of controls never diverge
- The seventies exclosures exceed controls by period 13
- The eighties exclosures exceed controls by period 119, which is Feb 1988
- The eighties exclosures (mostly) diverge from the seventies exclosures until period 121, which is April 1988. 

The eighties exclosures were implemented in Jan 1988. 

So,

- An almost immediate effect
- And no appreciable lag (mayyyybe 2 months but I wouldn't put any stake in that) in new exclosures matching old exclosures.

This is a lot of pairwise comparisons, but I think this may be the best way to do this? Because I don't think you can do factor interactions in a GAM, and even if you could it's a little mindbending how to set them up (because the treatments change). 

There's also a fair amount of treatment mess embedded in the "control"/"exclosure" designations. Various seed additions, plant removals, ant removals. None of which were thought to have an impact, but it seems like it would be good practice to *confirm*. 