---
title: "Small granivores originals"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rats <- read.csv(here::here("lore", "1994_longterm", "1994_data_complete.csv"), stringsAsFactors = F)
rat_plot_totals <- read.csv(here::here("lore", "1994_longterm", "1994_data_plot_totals.csv"), stringsAsFactors = F)
library(dplyr)
library(ggplot2)

```

```{r}

rat_plot_totals <- rat_plot_totals %>%
  mutate(plot_type = factor(plot_type),
         plot = factor(plot)) %>%
  mutate(oplot_type = ordered(plot_type),
         oplot = ordered(plot))
```


### Plot counts but plot not included as a predictor

```{r}
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))

sg_all<- filter(rat_plot_totals, type == "small_granivore")

ggplot(sg_all, aes(period, group = oplot,  nind, color = oplot_type)) +
  geom_line(size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) + 
  theme(legend.position = "top") + facet_wrap(vars(oplot_type))

library(mgcv)
sg.gam <- gam(nind ~ oplot_type + s(period, k = 50) + s(period, by = oplot_type, k = 50), family = "poisson", data = sg_all, method = "REML", control = gam.control(nthreads = 4))
summary(sg.gam)

gam.check(sg.gam)

sg.pdat <- make_pdat(sg_all, comparison_variable = "oplot_type") %>%
  mutate(plot_type = levels(sg_all$plot_type)[1])
sg.pred <- get_predicted_vals(sg.gam, sg.pdat)


#sg.pred <- add_exclosure_diff(sg.pred, sg.diff)

print(plot_link_pred(sg.pred, comparison_variable = "oplot_type"))

print(plot_fitted_pred(sg.pred, comparison_variable = "oplot_type"))


seventies_pred <- filter(sg.pred, grepl("orig", oplot_type))
print(plot_fitted_pred(seventies_pred, comparison_variable = "oplot_type"))

seventies_diff <- get_exclosure_diff(sg.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 1, comparison_level = 2)

print(plot_exclosure_diff(seventies_diff))

print(plot_fitted_pred(add_exclosure_diff(seventies_pred, seventies_diff), comparison_variable = "oplot_type"))


ggplot(filter(rat_plot_totals, type == "small_granivore", grepl("orig", plot_type)), aes(period, nind, color = oplot_type, group = plot)) +
  geom_line() +
  theme_bw() +
  scale_color_viridis_d(end = .8) 

```


### Including plot as a predictor

```{r}
source(here::here("lore", "1994_longterm", "gams_fxns_generalized.R"))

sg_all<- filter(rat_plot_totals, type == "small_granivore")

ggplot(sg_all, aes(period, group = oplot,  nind, color = oplot_type)) +
  geom_line(size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) + 
  theme(legend.position = "top") + facet_wrap(vars(oplot_type))

library(mgcv)
sg.p.gam <- gam(nind ~ oplot_type + oplot + s(period, k = 20) + s(period, by = oplot_type, k =20) + s(period, by = oplot, k= 20), family = "poisson", data = sg_all, method = "REML", control = gam.control(nthreads = 4))
summary(sg.p.gam)

gam.check(sg.p.gam)

sg.pdat <- make_pdat(sg_all, comparison_variable = "oplot_type", include_plot = T) %>%
  mutate(plot_type = levels(sg_all$plot_type)[1],
         oplot = levels(sg_all$oplot)[1])
sg.pred <- get_predicted_vals(sg.p.gam, sg.pdat)


#sg.pred <- add_exclosure_diff(sg.pred, sg.diff)

print(plot_link_pred(sg.pred, comparison_variable = "oplot_type"))

print(plot_fitted_pred(sg.pred, comparison_variable = "oplot_type"))


seventies_pred <- filter(sg.pred, grepl("orig", oplot_type))
print(plot_fitted_pred(seventies_pred, comparison_variable = "oplot_type"))

seventies_diff <- get_exclosure_diff(sg.p.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 1, comparison_level = 2)

print(plot_exclosure_diff(seventies_diff))

print(plot_fitted_pred(add_exclosure_diff(seventies_pred, seventies_diff), comparison_variable = "oplot_type"))

```
 
 This model behaves weirdly if:
 
 - you include `plot`, not `oplot`, as the parametric term. That then gave predictions where the *controls* exceed the *exclosures* in time periods where that's definitely not what the data are doing
 - you increase k to 30. That then fits wildly high numbers for the orig_exclosures.
 
I have not managed to increase k sufficiently to get out of "low p value" k-check land. High k for the ones that include plot takes a long time to run, plus there's the weirdly enormous values for `orig_exclosure` for k = 30. 


```{r}


seventies_diff %>%
  filter(diff_overlaps_zero) %>%
  filter(period == max(period)) %>%
  distinct()

eighties_pred <- filter(sg.pred, oplot_type %in% c("second_control", "second_exclosure"))

print(plot_fitted_pred(eighties_pred, comparison_variable = "oplot_type"))

eighties_diff <- get_exclosure_diff(sg.p.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 3, comparison_level = 4)

print(plot_exclosure_diff(eighties_diff))

eighties_pred <- add_exclosure_diff(eighties_pred, eighties_diff)
print(plot_fitted_pred(eighties_pred, comparison_variable = "oplot_type"))

max(eighties_pred$period [ which(eighties_pred$diff_overlaps_zero)])

controls_pred <- filter(sg.pred, oplot_type %in% c("second_control", "orig_control"))
print(plot_fitted_pred(controls_pred, comparison_variable = "oplot_type"))

controls_diff <- get_exclosure_diff(sg.p.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 1, comparison_level = 3)
print(plot_exclosure_diff(controls_diff))

controls_pred <- add_exclosure_diff(controls_pred, controls_diff)
print(plot_fitted_pred(controls_pred, comparison_variable = "oplot_type"))

ggplot(filter(sg_all, grepl("control", oplot_type)), aes(period, nind, group = plot, color = oplot_type)) +
  geom_line() +
  scale_color_viridis_d(end = .8) +
  theme_bw() + facet_wrap(vars(oplot_type)) +
  xlim(95, 150)

print(plot_fitted_pred(controls_pred, comparison_variable = "oplot_type") + xlim(95, 150))

exclosures_pred <- filter(sg.pred, oplot_type %in% c("second_exclosure", "orig_exclosure"))
print(plot_fitted_pred(exclosures_pred, comparison_variable = "oplot_type"))

exclosures_diff <- get_exclosure_diff(sg.p.gam, sg.pred, comparison_variable = "oplot_type", reference_level = 2, comparison_level = 4)
print(plot_exclosure_diff(exclosures_diff))

exclosures_pred <- add_exclosure_diff(exclosures_pred, exclosures_diff)
print(plot_fitted_pred(exclosures_pred, comparison_variable = "oplot_type"))

exclosures_pred %>%
  filter(period > 100) %>%
  filter(diff_overlaps_zero) %>%
  select(period) %>%
  distinct() %>%
  filter(period == min(period))



```
