---
title: "1981 analysis with GAMs"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 4, fig.height=3)
rats <- read.csv(here::here("lore", "1981_competition", "1981_data_complete.csv"), stringsAsFactors = F)
rat_totals <- read.csv(here::here("lore", "1981_competition", "1981_data_statevars.csv"), stringsAsFactors = F)
rat_plot_totals <- read.csv(here::here("lore", "1981_competition", "1981_data_plot_totals.csv"), stringsAsFactors = F)
library(dplyr)
library(mgcv)
library(ggplot2)
source(here::here("lore", "1981_competition", "gams_fxns.R"))



rat_totals <- rat_totals %>%
  mutate(krat_treatment = ifelse(brown_trtmnt == "dipo_present", "control", "exclosure")) %>%
  mutate(okrat_treatment = ordered(krat_treatment))

```

```{r, echo = F, fig.dim = c(6,5)}

ggplot(filter(rat_totals, type != "other"), aes(period, nind, color = okrat_treatment)) +
  geom_line() +
  geom_point()+
  theme_bw() +
  facet_wrap(vars(type), ncol = 1, scales = "free_y") +
  geom_vline(xintercept = 3.5) +
  geom_hline(yintercept = 0) +
  scale_colour_viridis_d(end = .8) +
  theme(legend.position = "top")

```

```{r sg}
sg <- filter(rat_totals, type == "small_granivore")

print(plot_orig_data(sg))

sg.gam <- gam(nind ~ okrat_treatment + s(period) + s(period, by = okrat_treatment), data = sg, family = poisson, method = "REML")

summary(sg.gam)

sg.pdat <- make_pdat(sg)
sg.pred <- get_predicted_vals(sg.gam, sg.pdat)

sg.diff <- get_exclosure_diff(sg.gam, sg.pdat)

print(plot_exclosure_diff(sg.diff))

sg.pred <- add_exclosure_diff(sg.pred, sg.diff)

print(plot_link_pred(sg.pred))

print(plot_fitted_pred(sg.pred))

```

The small granivore timeseries start out separate (before the treatments are even implemented). Later they diverge starting around period `r trunc(min(filter(sg.pred, !diff_overlaps_zero, period > 4)$period))` and do not come back together.

```{r dipo}
dipo <- filter(rat_totals, type == "dipo")

print(plot_orig_data(dipo))

dipo.gam <- gam(nind ~ okrat_treatment + s(period) + s(period, by = okrat_treatment), data = dipo, family = poisson, method = "REML")

summary(dipo.gam)

dipo.pdat <- make_pdat(dipo)
dipo.pred <- get_predicted_vals(dipo.gam, dipo.pdat)

dipo.diff <- get_exclosure_diff(dipo.gam, dipo.pdat)

print(plot_exclosure_diff(dipo.diff))

dipo.pred <- add_exclosure_diff(dipo.pred, dipo.diff)

print(plot_link_pred(dipo.pred))

print(plot_fitted_pred(dipo.pred))


```

The dipo timeseries diverge starting around period `r trunc(min(filter(dipo.pred, !diff_overlaps_zero)$period))` and do not come back together.



```{r small_omniv}
small_omniv <- filter(rat_totals, type == "small_omnivore")

print(plot_orig_data(small_omniv))

small_omniv.gam <- gam(nind ~ okrat_treatment + s(period) + s(period, by = okrat_treatment), data = small_omniv, family = poisson, method = "REML")

summary(small_omniv.gam)

small_omniv.pdat <- make_pdat(small_omniv)
small_omniv.pred <- get_predicted_vals(small_omniv.gam, small_omniv.pdat)

small_omniv.diff <- get_exclosure_diff(small_omniv.gam, small_omniv.pdat)

print(plot_exclosure_diff(small_omniv.diff))

small_omniv.pred <- add_exclosure_diff(small_omniv.pred, small_omniv.diff)

print(plot_link_pred(small_omniv.pred))

print(plot_fitted_pred(small_omniv.pred))


```

```{r, fig.dim = c(4, 8)}

sg_res <- plot_fitted_pred(sg.pred)
dipo_res <- plot_fitted_pred(dipo.pred)
so_res <- plot_fitted_pred(small_omniv.pred)

gridExtra::grid.arrange(grobs = list(sg_res, dipo_res, so_res), ncol = 1)
```