---
title: "New plots with 81 data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rats <- read.csv(here::here("lore", "1981_competition", "1981_data_complete.csv"), stringsAsFactors = F)
rat_totals <- read.csv(here::here("lore", "1981_competition", "1981_data_statevars.csv"), stringsAsFactors = F)
library(dplyr)
library(ggplot2)
library(gratia)
```

```{r, echo = F}

rat_totals <- rat_totals %>%
  mutate(krat_treatment = ifelse(brown_trtmnt == "dipo_present", "control", "exclosure")) %>%
  mutate(okrat_treatment = ordered(krat_treatment))

ggplot(filter(rat_totals, type != "other"), aes(period, nind, color = krat_treatment)) +
  geom_line() +
  geom_point()+
  theme_bw() +
  facet_wrap(vars(type), ncol = 1, scales = "free_y") +
  geom_vline(xintercept = 3.5) +
  geom_hline(yintercept = 0) +
  scale_colour_viridis_d(end = .8)

```
```{r}


load_mgcv()


sg <- filter(rat_totals, type == "small_granivore")

n_gam <- gam(nind ~  okrat_treatment + s(period) + s(period, by = okrat_treatment), data = sg, method = "REML", family = "poisson")

n_gam_fitted <- add_fitted(sg, n_gam, value = "fitted")

ggplot(n_gam_fitted, aes(period, nind, color = krat_treatment)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  geom_line(aes(period, fitted, color = krat_treatment), size = 2) +
  scale_color_viridis_d(end = .8)

fitted_ci <- function(gam_obj, ndraws = 500, df, seed = 1977) {
  
  sampled_vals <- fitted_samples(gam_obj, n = ndraws, newdata = df, seed = seed)
  
  sampled_vals <- sampled_vals %>%
    group_by(row) %>%
    summarize(
      meanfit = mean(fitted),
      lowerfit = quantile(fitted, probs = .025),
      upperfit= quantile(fitted, probs = .975)
    ) %>%
    ungroup()
  
  df <- df %>%
    mutate(row = dplyr::row_number()) %>%
    left_join(sampled_vals)
  
  df  
}

n_gam_ci_manual <- fitted_ci(n_gam, df= sg)


ggplot(n_gam_ci_manual, aes(period, nind, color = krat_treatment)) +
  geom_line() +
  geom_line(aes(period, meanfit)) +
  geom_ribbon(aes(period, ymin = lowerfit, ymax = upperfit, fill = krat_treatment), alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") +
  scale_color_viridis_d(end = .8) +
  scale_fill_viridis_d(end = .8)

```

The `difference_smooths` function in `gratia` doesn't work for ordered smooths. Working from Gavin's post...

```{r}

plot(n_gam, shade = T, pages = 1, scale = 0, seWithMean = T)
```

This may have directly given you what you want - the difference between the control and the exclosure smooths is the plot on the right. 

We can apparently interpret the signifiance relatively usually? 

```{r}
summary(n_gam)

```
I.e. the exclosure is significant. 

We could use a not ordered factor to estimate the parametric effect if we want an easy-to-interpret mean?

```{r}
n_gam2 <- n_gam <- gam(nind ~  krat_treatment + s(period) + s(period, by = okrat_treatment), data = sg, method = "REML", family = "poisson")

summary(n_gam2)
```

Can I get the difference smooth?

```{r}
pdat <- data.frame(period = seq(min(sg$period), max(sg$period), length.out = 100),
                   krat_treatment = sg$krat_treatment[1],
                   okrat_treatment= sg$okrat_treatment[1])


xp <- predict(n_gam, newdata = pdat, type = 'lpmatrix')

# cols for the difference smooth
diff_cols <- grepl("okrat_treatmentexclosure", colnames(xp))

X <- xp
# Zero out the columns from the xp matrix that were for the other smooth

X[, !diff_cols] <- 0

# And zero out ones that aren't smooths (i.e. the parametric cols)
X[ ,!grepl("^s\\(", colnames(xp))] <- 0

# Use matrix multiplication of the X matrix (the modified Xp matrix) with the model coefficients from m to MULTIPLY Xp by the coefficients and SUM row wise in one step

dsmooth <- X %*% coef(n_gam)

plot(dsmooth)

plot(n_gam, shade = T, pages = 1, scale = 0, seWithMean = T)

# Might be able to do this directly using `mgcv::predict.gam`

diff_direct <- predict(n_gam, newdata = pdat, type = "link", se.fit = T, terms = "s(period):okrat_treatmentexclosure")

plot(diff_direct$fit)
```

These methods are all giving somewhat different results and I don't like that. 