---
title: "GAMs for ratios"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.dim = c(6, 3))
library(soar)
library(ggplot2)
library(dplyr)
library(mgcv)
library(gratia)
theme_set(theme_bw())
```

For two of the major pieces of this analysis, we are primarily interested in how certain ratios have changed over time: the ratio of total energy use on treatment:control plots, and the amount of compensation happening from non-Dipos on treatment relative to control plots. 

This is complicated somewhat by the experimental setup: we have 3-5 of each type of plot, and there's variability between plots within a treatment. 

The naive way to handle this is to take the treatment-level mean for each time step, compute the ratio, and analyze that. This is analogous to what was done in Ernest and Brown (2001). (They summed across equal numbers of plots).

```{r}

treatl <- get_treatment_means() %>%
  mutate(fplottype = as.factor(plot_type))

just_controls <- filter(treatl, oplottype == "CC")

era_grid <-   facet_grid(cols = vars(oera), space = "free", scales = "free_x")

totale_ratio <- treatl %>%
  filter(oplottype != "CC") %>%
  select(period, censusdate, era, oera, plot_type, oplottype, total_e, fplottype) %>%
  left_join(select(just_controls, period, total_e) %>% rename(total_e_c = total_e)) %>%
  mutate(total_e_rat = total_e / total_e_c) %>%
  group_by(oplottype) %>%
  mutate(total_e_rat_ma = maopts(total_e_rat)) %>%
  ungroup()

ggplot(totale_ratio, aes(censusdate, total_e_rat_ma, color = oplottype)) +
  geom_line() +
  ggtitle("Total E ratio, from treatment means") +
  era_grid

compensation <- treatl %>%
  filter(oplottype != "CC") %>%
  select(period, censusdate, era, oera, plot_type, oplottype, smgran_e, fplottype) %>%
  left_join(select(just_controls, period, dipo_e, smgran_e) %>% rename(dipo_e_c = dipo_e, smgran_e_c = smgran_e)) %>%
  mutate(smgran_increase = smgran_e - smgran_e_c) %>%
  mutate(smgran_comp = smgran_increase / dipo_e_c)%>%
  group_by(oplottype) %>%
  mutate(smgran_comp_ma = maopts(smgran_comp)) %>%
  ungroup()


ggplot(compensation, aes(censusdate, smgran_comp_ma, color = oplottype)) +
  geom_line() +
  ggtitle("Small granivore compensation, from treatment means") +
  era_grid

```

You could then fit a model of the general structure `response ~ oera * oplottype, cor = corAR1` and investigate contrasts between, for example, compensation on EEs in c vs b. 

However, if possible, it would be nice to incroporate the plot-level variability into our estimates of these quantities.

Following is me doing this using GAMs. 

### GAMs for plot level variability, total e

```{r, echo = T}

plotl <- get_plot_totals() %>%
  mutate(fplottype = as.factor(plot_type),
         fera = as.factor(era))


totale_gam <- gam(total_e ~ fplottype + s(period, by = fplottype) + s(plot, bs = "re"), data = plotl, family = "tw")

plotl_pdat <- plotl %>%
  select(censusdate, period, fera, fplottype, oera) %>%
  distinct()

totale_gam_pred <- plotl_pdat %>%
  add_fitted(totale_gam, exclude = "s(plot)", newdata.guaranteed =T)

ggplot(totale_gam_pred, aes(censusdate, .value, color = fplottype)) +
  geom_line()

totale_gam_sim <- fitted_samples(totale_gam, n = 100, newdata = plotl_pdat, exclude = "s(plot)", newdata.guaranteed = T, seed = 1977) %>%
  left_join(mutate(plotl_pdat, row = dplyr::row_number()))

controls_gam_sim <- filter(totale_gam_sim, fplottype == "CC")

ratio_gam_sim <- totale_gam_sim %>%
  filter(fplottype != "CC") %>%
  left_join(rename(select(controls_gam_sim, draw, fitted, period, fera, oera), fitted_c = fitted)) %>%
  mutate(totale_rat = fitted / fitted_c) %>%
  group_by(censusdate, period, fera, oera, fplottype) %>%
  summarize(totale_rat_mean = mean(totale_rat),
         totale_rat_lwr = quantile(totale_rat, .025),
         totale_rat_upr = quantile(totale_rat, .975)) %>%
  ungroup() 

ggplot(ratio_gam_sim, aes(censusdate, totale_rat_mean, color  = fplottype, fill = fplottype)) +
  geom_line() +
  geom_ribbon(aes(ymin  = totale_rat_lwr,
                  ymax = totale_rat_upr), alpha = .3) +
  ggtitle("Total energy ratio via GAM sims") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  era_grid


```

These estimates incorporate among-plot variability in the estimates. Note that these 95% intervals are not **simultaneous** intervals, because the calculation for simultaneous intervals comparing factor levels doesn't work if there are ratios involved. 

Compared to the means, you end up in a very similar place:

```{r}


ggplot(ratio_gam_sim, aes(censusdate, totale_rat_mean, color  = fplottype, fill = fplottype)) +
  geom_line() +
  geom_ribbon(aes(ymin  = totale_rat_lwr,
                  ymax = totale_rat_upr), alpha = .3) +
  ggtitle("Total energy ratio via GAM sims compared to means") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  geom_line(data = totale_ratio, aes(censusdate, total_e_rat_ma, color = oplottype), linetype = 2) +
  facet_grid(rows = vars(fplottype), cols = vars(oera), space = "free_x", scales = "free_x")

```

### Compensation
```{r, echo = T}

comp_data <- plotl %>%
  select(period, censusdate, era, oera, plot, plot_type, fplottype, oplottype, fera, dipo_e, smgran_e) %>%
  tidyr::pivot_longer(c("dipo_e", "smgran_e"), names_to = "rod_group", values_to = "energy") %>%
  mutate(rod_group = as.factor(rod_group),
         rod_group_trt = paste0(rod_group, "_", plot_type)) %>%
  mutate(rod_group_trt = as.factor(rod_group_trt)) %>%
  filter(rod_group_trt %in% c("dipo_e_CC", "smgran_e_CC", "smgran_e_EE", "smgran_e_CE", "smgran_e_EC")) 


comp_gam <- gam(energy ~ fplottype + rod_group + s(period, by = fplottype) + s(period, by = rod_group) + s(plot, bs = "re"), data = comp_data, family = "tw")

comp_pdat <- comp_data %>%
  select(censusdate, period, fera, fplottype, oera, rod_group) %>%
  distinct()

comp_gam_pred <- comp_pdat %>%
  add_fitted(comp_gam, exclude = "s(plot)", newdata.guaranteed =T)

ggplot(comp_gam_pred, aes(censusdate, .value, color = fplottype)) +
  geom_line() +
  facet_wrap(vars(rod_group))

comp_gam_sim <- fitted_samples(comp_gam, n = 100, newdata = comp_pdat, exclude = "s(plot)", newdata.guaranteed = T, seed = 1977) %>%
  left_join(mutate(comp_pdat, row = dplyr::row_number()))

controls_comp_gam_sim <- filter(comp_gam_sim, fplottype == "CC") %>%
  select(-row) %>%
  tidyr::pivot_wider(names_from = rod_group, values_from = fitted) %>%
  rename(dipo_e_c = dipo_e,
         smgran_e_c = smgran_e) %>%
  select(-fplottype)

smgran_comp_gam_sim <- comp_gam_sim %>%
  filter(fplottype != "CC") %>%
  select(-row) %>%
  tidyr::pivot_wider(names_from = rod_group, values_from = fitted) %>% 
  left_join(controls_comp_gam_sim) %>%
  mutate(smgran_increase = smgran_e - smgran_e_c) %>%
  mutate(smgran_comp = smgran_increase/ dipo_e_c) %>%
  group_by(censusdate, period, fera, oera, fplottype) %>%
  summarize(smgran_comp_mean = mean(smgran_comp),
         smgran_comp_lwr = quantile(smgran_comp, .025),
         smgran_comp_upr = quantile(smgran_comp, .975)) %>%
  ungroup() 

ggplot(smgran_comp_gam_sim, aes(censusdate, smgran_comp_mean, color  = fplottype, fill = fplottype)) +
  geom_line() +
  geom_ribbon(aes(ymin  = smgran_comp_lwr,
                  ymax = smgran_comp_upr), alpha = .3) +
  ggtitle("Compensation via GAM sims") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  era_grid




ggplot(smgran_comp_gam_sim, aes(censusdate, smgran_comp_mean, color  = fplottype, fill = fplottype)) +
  geom_line() +
  geom_ribbon(aes(ymin  = smgran_comp_lwr,
                  ymax = smgran_comp_upr), alpha = .3) +
  ggtitle("Compensation via GAM sims compared to means") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  geom_line(data =compensation, aes(censusdate, smgran_comp_ma, color = fplottype), linetype = 2) +
  facet_grid(rows = vars(fplottype), cols = vars(oera), space = "free_x", scales = "free_x")

```

Again, you end up in a very similar place to using just the means.

However, note that in both the compensation approaches, there's the possibility that the treatment small granivores are not "significantly" greater than the controls, which casts some doubt on the "significance" of the compensation effect. Currently my best thought for how to represent this is to fit `gam(smgran_e ~ fplottype + s(period, by = fplottype) + s(plot, bs = "re"))`, identify the periods of time when the different treatment smooths overlap the control smooth, and flag those as periods when the "significance" of the compensatory effect is suspect. 

## Contrast models


Whether we use the GAM sim method or just the treatment means, we would ultimately like a - general - estimate of the compensation and total energy values for the different "eras". 

Suggestion: Use a generalized least squares. This will fit the ratio as Gaussian, which may not be totally correct...but at least it is a continuous value that can be positive or negative. It will allow for an autocorrelation term. 

### Contrasts on treatment means


```{r}

library(nlme)
library(emmeans)


totale_mean_gls <- gls(total_e_rat ~ plot_type * era,  correlation = corCAR1(form = ~ period | plot_type), data = totale_ratio)

totale_mean_gls_emmeans <- emmeans(totale_mean_gls, specs = ~ era | plot_type)

totale_mean_lm <- lm(total_e_rat ~ plot_type * era, data = totale_ratio)

totale_mean_lm_emmeans <- emmeans(totale_mean_lm, specs = ~ era | plot_type)


anova(totale_mean_gls, totale_mean_lm)

totale_mean_pairs <- bind_rows(gls = as.data.frame(pairs(totale_mean_gls_emmeans)),
                              lm = as.data.frame(pairs(totale_mean_lm_emmeans)),
                              .id = "mod")

ggplot(totale_mean_pairs, aes(contrast, estimate, shape = mod, color = p.value < .05)) +
  geom_jitter() +
  facet_wrap(vars(plot_type)) +
  theme(axis.text.x = element_text(angle = 90))

totale_mean_pred <- bind_rows(gls = as.data.frame(totale_mean_gls_emmeans),
                              lm = as.data.frame(totale_mean_lm_emmeans),
                              .id = "mod")

ggplot(filter(totale_mean_pred, mod == "gls"), aes(era, emmean, color = mod)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL)) +
  facet_wrap(vars(plot_type))

```

```{r}

comp_mean_gls <- gls(smgran_comp ~ plot_type * era,  correlation = corCAR1(form = ~ period | plot_type), data = compensation)

comp_mean_gls_emmeans <- emmeans(comp_mean_gls, specs = ~ era | plot_type)

comp_mean_lm <- lm(smgran_comp ~ plot_type * era, data = compensation)

comp_mean_lm_emmeans <- emmeans(comp_mean_lm, specs = ~ era | plot_type)


anova(comp_mean_gls, comp_mean_lm)

comp_mean_pairs <- bind_rows(gls = as.data.frame(pairs(comp_mean_gls_emmeans)),
                              lm = as.data.frame(pairs(comp_mean_lm_emmeans)),
                              .id = "mod")

ggplot(comp_mean_pairs, aes(contrast, estimate, color = p.value < .05, shape = mod)) +
  geom_jitter() +
  facet_wrap(vars(plot_type))+
  theme(axis.text.x = element_text(angle = 90))

comp_mean_pred <- bind_rows(gls = as.data.frame(comp_mean_gls_emmeans),
                              lm = as.data.frame(comp_mean_lm_emmeans),
                              .id = "mod")
ggplot(filter(comp_mean_pred, mod == "gls"), aes(era, emmean, color = mod)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL)) +
  facet_wrap(vars(plot_type))

```


## Contrasts on gam sims

There's something weird about the GLS on the gam sims. It has enormous CIs and is attributing essentially no change to era. 

I also have a hard time fitting multiple grouping factors for the correlation in the GLS, and so went with the means (and not taking into account the draws). 

For now I suggest using the GLS on the means.


<!-- ```{r} -->

<!-- ratio_gam_sim <- ratio_gam_sim %>% -->
<!--   mutate(plot_type = as.character(fplottype), -->
<!--          era = as.character(fera)) -->

<!-- totale_sim_gls <- gls(totale_rat_mean ~ plot_type * era,  correlation = corCAR1(form = ~ period | plot_type), data = ratio_gam_sim) -->

<!-- totale_sim_gls_emmeans <- emmeans(totale_sim_gls, specs = ~ era | plot_type) -->

<!-- totale_sim_lm <- lm(totale_rat_mean ~ plot_type * era, data = ratio_gam_sim) -->

<!-- totale_sim_lm_emmeans <- emmeans(totale_sim_lm, specs = ~ era | plot_type) -->

<!-- anova(totale_sim_gls, totale_sim_lm) -->

<!-- totale_sim_pairs <- bind_rows(gls = as.data.frame(pairs(totale_sim_gls_emmeans)), -->
<!--                               lm = as.data.frame(pairs(totale_sim_lm_emmeans)), -->
<!--                               .id = "mod") -->

<!-- ggplot(totale_sim_pairs, aes(contrast, estimate, shape = mod, color = p.value < .05)) + -->
<!--   geom_jitter() + -->
<!--   facet_wrap(vars(plot_type)) + -->
<!--   theme(axis.text.x = element_text(angle = 90)) -->

<!-- totale_sim_pred <- bind_rows(gls = as.data.frame(totale_sim_gls_emmeans), -->
<!--                               lm = as.data.frame(totale_sim_lm_emmeans), -->
<!--                               .id = "mod") -->

<!-- ggplot(filter(totale_sim_pred), aes(era, emmean, color = mod)) + -->
<!--   geom_point() + -->
<!--   #geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL)) + -->
<!--   facet_wrap(vars(plot_type)) -->




<!-- ``` -->


<!-- ```{r} -->

<!-- smgran_comp_gam_sim <- smgran_comp_gam_sim %>% -->
<!--   mutate(plot_type = as.character(fplottype), -->
<!--          era = as.character(fera)) -->

<!-- comp_sim_gls <- gls(smgran_comp_mean ~ plot_type * era,  correlation = corCAR1(form = ~ period | plot_type), data = smgran_comp_gam_sim) -->

<!-- comp_sim_gls_emmeans <- emmeans(comp_sim_gls, specs = ~ era | plot_type) -->

<!-- comp_sim_lm <- lm(smgran_comp_mean ~ plot_type * era, data = smgran_comp_gam_sim) -->

<!-- comp_sim_lm_emmeans <- emmeans(comp_sim_lm, specs = ~ era | plot_type) -->

<!-- anova(comp_sim_gls, comp_sim_lm) -->

<!-- comp_sim_pairs <- bind_rows(gls = as.data.frame(pairs(comp_sim_gls_emmeans)), -->
<!--                               lm = as.data.frame(pairs(comp_sim_lm_emmeans)), -->
<!--                               .id = "mod") -->

<!-- ggplot(comp_sim_pairs, aes(contrast, estimate, shape = mod, color = p.value < .05)) + -->
<!--   geom_jitter() + -->
<!--   facet_wrap(vars(plot_type)) + -->
<!--   theme(axis.text.x = element_text(angle = 90)) -->

<!-- comp_sim_pred <- bind_rows(gls = as.data.frame(comp_sim_gls_emmeans), -->
<!--                               lm = as.data.frame(comp_sim_lm_emmeans), -->
<!--                               .id = "mod") -->

<!-- ggplot(filter(comp_sim_pred), aes(era, emmean, color = mod)) + -->
<!--   geom_point() + -->
<!--   #geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL)) + -->
<!--   facet_wrap(vars(plot_type)) -->




<!-- ``` -->
